
This file provides code to perform descriptive analyses. This includes creating the Table 1, environmental variable correlation analyses, examining differences in distributions across study centers, and assessment of continuous variable distributions.


```{r}
library(patchwork)
library(dplyr)
library(ggplot2)
library(tableone)
```


#0. Load data

Load datasets that have already been cleaned and include all relevant feature engineering.

u1_df: 
- df for NAKO baseline
- complete case for baseline COPD, baseline asthma, covariates, and exposures

u2_df_asthma: 
- df for the 5-year NAKO Follow-up
- individuals that did not have asthma at baseline
- complete case for covariates, incident asthma, and exposures

u2_df_copd: 
- df for the 5-year NAKO Follow-up
- individuals that did not have COPD at baseline
- complete case for covariates, incident COPD, and exposures
```{r}
u1_df        <- readRDS("")
u2_df_asthma <- readRDS("")
u2_df_copd   <- readRDS("")
```


#1. Creating a table 1

functions
```{r}
summary_continuous <- function(df, var_name, var) {
  
  # Detect whether df name starts with "u1" or "u2"
  df_name <- deparse(substitute(df))
  
  if (startsWith(df_name, "u1")) {
    asthma_var <- "u1_d_an_lung_2"
    copd_var   <- "u1_d_an_lung_1"
  } else if (startsWith(df_name, "u2")) {
    asthma_var <- "u2_dt_an_lung_2_fu"
    copd_var   <- "u2_dt_an_lung_1_fu"
  } else {
    stop("Dataframe name must start with 'u1' or 'u2'")
  }
  
  get_mean_sd <- function(x) {
  m <- mean(x, na.rm = TRUE)
  iqr <- IQR(x, na.rm = TRUE)
  sprintf("%.2f (%.2f)", m, iqr)
  }
  
  tibble::tibble(
    #Variable = paste0(var_name, ", mean (sd) "),
    Variable = paste0(var_name, ", mean (IQR)"),
    #Overall  = get_mean_sd(df[[var]]),
    #Overall   = get_mean_sd(df[[var]]),
    Asthma   = get_mean_sd(df %>% filter(.data[[asthma_var]] == 1) %>% pull(var)),
    No_Asthma= get_mean_sd(df %>% filter(.data[[asthma_var]] == 0) %>% pull(var)),
    COPD     = get_mean_sd(df %>% filter(.data[[copd_var]]   == 1) %>% pull(var)),
    No_COPD  = get_mean_sd(df %>% filter(.data[[copd_var]]   == 0) %>% pull(var))
  )
}

summary_categorical <- function(df, var_name, var) {

  # Detect df name (u1 vs u2)
  df_name <- deparse(substitute(df))

  if (startsWith(df_name, "u1")) {
    asthma_var <- "u1_d_an_lung_2"
    copd_var   <- "u1_d_an_lung_1"
  } else if (startsWith(df_name, "u2")) {
    asthma_var <- "u2_dt_an_lung_2_fu"
    copd_var   <- "u2_dt_an_lung_1_fu"
  } else {
    stop("Dataframe name must start with 'u1' or 'u2'")
  }

  # Get full categories but EXCLUDE NA
  categories <- names(table(df[[var]], useNA = "no"))

  make_tab <- function(data, var) {
    vec <- factor(data[[var]], levels = categories)
    tab <- table(vec)
    pct <- prop.table(tab) * 100
    sprintf("%d (%.2f%%)", as.integer(tab), pct)
  }

  tibble::tibble(
    Variable  = paste0(var_name, ": ", categories, ", n (%)"),
    Overall   = make_tab(df, var),
    Asthma    = make_tab(dplyr::filter(df, .data[[asthma_var]] == 1), var),
    No_Asthma = make_tab(dplyr::filter(df, .data[[asthma_var]] == 0), var),
    COPD      = make_tab(dplyr::filter(df, .data[[copd_var]]   == 1), var),
    No_COPD   = make_tab(dplyr::filter(df, .data[[copd_var]]   == 0), var)
  )
}
```

1.1 create table 1 for baseline
```{r}
u1_table1 <- dplyr::bind_rows(
  summary_continuous(u1_df, "Age", "u1_basist_age"),
  summary_categorical(u1_df, "Sex", "u1_basist_sex"),
  summary_categorical(u1_df, "Income below median", "u1_at_ses_inc_median"),
  summary_categorical(u1_df, "Migration status", "u1_at_ses_mig_status"),
  summary_categorical(u1_df, "Education", "u1_at_ses_isced11_3level"),
  summary_categorical(u1_df, "Smoking status", "u1_at_smok_stat_noint"),
  summary_continuous(u1_df, "BMI", "u1_at_anthro_bmi"),
  #physical activity with quartiles does not add any information to table 1--all 25%
  summary_continuous(u1_df, "Physical Activity", "u1_at_gpaq_ptotalday"),
  summary_categorical(u2_df_asthma, "Asthma", "u1_d_an_lung_2"),
  summary_categorical(u2_df_asthma, "COPD", "u1_d_an_lung_1"),
  summary_categorical(u1_df, "Occupational exposure", "u1_d_bb1"),
  summary_continuous(u1_df, "Mean annual temperature", "u1_gt_t_mean_hmg_g"),
  summary_continuous(u1_df, "Standard deviation annual temperature", "u1_gt_t_sd_hmg_g"),
  summary_continuous(u1_df, "O3", "u1_gt_o3_exp_g"),
  summary_continuous(u1_df, "PM2.5", "u1_gt_pm25_exp_g"),
  summary_continuous(u1_df, "PM10", "u1_gt_pm10_exp_g"),
  summary_continuous(u1_df, "NO2", "u1_gt_no2_exp_g"),
  summary_continuous(u1_df, "NDVI", "u1_gt_ndvi_1000_exp_g_15"),
  summary_continuous(u1_df, "GISD", "u1_gt_gisd_rki_m"),
)

write.csv(u1_table1, "", row.names = FALSE)
```

1.2 create table 1 for 5-year follow-up-asthma
```{r}
u2_asthma_table1 <- dplyr::bind_rows(
  #only need age for follow-up
  summary_continuous(u2_df_asthma, "Age", "u1_basist_age"),
  summary_categorical(u2_df_asthma, "Age category", "u1_basist_age_category"),
  #don't need to have sex for both--limited changes expected
  summary_categorical(u2_df_asthma, "Sex", "u1_basist_sex"),
  summary_categorical(u2_df_asthma, "Migration status", "u1_at_ses_mig_status"),
  summary_categorical(u2_df_asthma, "Education", "u1_at_ses_isced11_3level"),
  summary_categorical(u2_df_asthma, "Smoking status", "u1_at_smok_stat_noint"),
  #bmi only for follow-up
  summary_categorical(u2_df_asthma, "BMI", "u1_at_anthro_bmi_category"),
  #updated physical activity
  summary_continuous(u2_df_asthma, "Physical Activity", "u1_at_gpaq_ptotalday"),
  summary_categorical(u2_df_asthma, "Prevalent Asthma", "u1_d_an_lung_2"),
  summary_categorical(u2_df_asthma, "Prevalent COPD", "u1_d_an_lung_1"),
  summary_categorical(u2_df_asthma, "Incident Asthma", "u2_dt_an_lung_2_fu"),
  summary_categorical(u2_df_asthma, "Incident COPD", "u2_dt_an_lung_1_fu"),
  summary_categorical(u2_df_asthma, "Occupational exposure", "u1_d_bb1"),
  summary_continuous(u2_df_asthma, "Mean annual temperature", "u1_gt_t_mean_hmg_g"),
  summary_continuous(u2_df_asthma, "Mean winter temperature", "u1_gt_t_win_mean_hmg_g"),
  summary_continuous(u2_df_asthma, "Mean spring temperature", "u1_gt_t_spr_mean_hmg_g"),
  summary_continuous(u2_df_asthma, "Mean summer temperature", "u1_gt_t_smr_mean_hmg_g"),
  summary_continuous(u2_df_asthma, "Mean fall temperature", "u1_gt_t_fal_mean_hmg_g"),
  summary_continuous(u2_df_asthma, "PM2.5", "u1_gt_pm25_exp_g"),
  summary_continuous(u2_df_asthma, "PM10", "u1_gt_pm10_exp_g"),
  summary_continuous(u2_df_asthma, "GISD", "u1_gt_gisd_rki_m"),
  summary_continuous(u2_df_asthma, "NO2", "u1_gt_no2_exp_g"),
  summary_continuous(u2_df_asthma, "Population Desnity", "u1_gt_pop_d_ink_a_18"),
  summary_categorical(u2_df_asthma, "Degree of Urbanization", "u1_gt_dgu_eu_m"),
  summary_categorical(u2_df_asthma, "Study center", "u1_basist_uort_combined")
)

write.csv(u2_asthma_table1, "", row.names = FALSE)
```

1.3 create table 1 for 5-year follow-up-COPD
```{r}
u2_copd_table1 <- dplyr::bind_rows(
  # only need age for follow-up
  summary_continuous(u2_df_copd, "Age", "u1_basist_age"),
  summary_categorical(u2_df_copd, "Age", "u1_basist_age_category"),
  # don't need to have sex for both--limited changes expected
  summary_categorical(u2_df_copd, "Sex", "u1_basist_sex"),
  summary_categorical(u2_df_copd, "Migration status", "u1_at_ses_mig_status"),
  summary_categorical(u2_df_copd, "Education", "u1_at_ses_isced11_3level"),
  summary_categorical(u2_df_copd, "Smoking status", "u1_at_smok_stat_noint"),
  # bmi only for follow-up
  summary_categorical(u2_df_copd, "BMI", "u1_at_anthro_bmi_category"),
  # updated physical activity
  summary_continuous(u2_df_copd, "Physical Activity", "u1_at_gpaq_ptotalday"),
  summary_categorical(u2_df_copd, "Prevalent Asthma", "u1_d_an_lung_2"),
  summary_categorical(u2_df_copd, "Prevalent COPD", "u1_d_an_lung_1"),
  summary_categorical(u2_df_copd, "Incident Asthma", "u2_dt_an_lung_2_fu"),
  summary_categorical(u2_df_copd, "Incident COPD", "u2_dt_an_lung_1_fu"),
  summary_categorical(u2_df_copd, "Occupational exposure", "u1_d_bb1"),
  summary_continuous(u2_df_copd, "Mean annual temperature", "u1_gt_t_mean_hmg_g"),
  summary_continuous(u2_df_copd, "Mean winter temperature", "u1_gt_t_win_mean_hmg_g"),
  summary_continuous(u2_df_copd, "Mean spring temperature", "u1_gt_t_spr_mean_hmg_g"),
  summary_continuous(u2_df_copd, "Mean summer temperature", "u1_gt_t_smr_mean_hmg_g"),
  summary_continuous(u2_df_copd, "Mean fall temperature", "u1_gt_t_fal_mean_hmg_g"),
  summary_continuous(u2_df_copd, "PM2.5", "u1_gt_pm25_exp_g"),
  summary_continuous(u2_df_copd, "PM10", "u1_gt_pm10_exp_g"),
  summary_continuous(u2_df_copd, "NO2", "u1_gt_no2_exp_g"),
  summary_continuous(u2_df_copd, "GISD", "u1_gt_gisd_rki_m"),
  summary_continuous(u2_df_copd, "Population Desnity", "u1_gt_pop_d_ink_a_18"),
  summary_categorical(u2_df_copd, "Degree of Urbanization", "u1_gt_dgu_eu_m"),
  summary_categorical(u2_df_copd, "Study center", "u1_basist_uort_combined")
)

write.csv(u2_copd_table1, "", row.names = FALSE)
```



#2. correlation analyses of environmental variables (Figure S3)

environmental variables of interest
```{r}
env_vars <- c( "u1_gt_t_mean_hmg_g", "u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g", "u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g", "u1_gt_no2_exp_g", "u1_gt_gisd_rki_m", "u1_gt_pop_d_ink_a_18")

env_labels <- c("Annual\nTemperature", "Fall\nTemperature", "Winter\nTemperature", "Spring\nTemperature", "Summer\nTemperature", "PM2.5", "PM10", "NO2", "GISD", "Population\nDensity")
```

histograms to check for normality; mostly normal except GISD and population density
-> use Spearman
```{r}
env_vars <- c(
  "u1_gt_t_mean_hmg_g",
  "u1_gt_t_win_mean_hmg_g",
  "u1_gt_t_fal_mean_hmg_g",
  "u1_gt_t_spr_mean_hmg_g",
  "u1_gt_t_smr_mean_hmg_g",
  "u1_gt_pm25_exp_g",
  "u1_gt_pm10_exp_g",
  "u1_gt_no2_exp_g",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18"
)

par(mfrow = c(3, 4))  # 3 rows, 4 columns

for (var in env_vars) {
  hist(
    u2_df_asthma[[var]],
    main = var,
    xlab = "",
    col = "lightblue",
    border = "white"
  )
}

par(mfrow = c(1, 1))  # reset plotting layout
```

function to get correlation matrix
```{r}
env <- u2_df_asthma[, c(env_vars)]

# Correlation matrix
cor_mat <- suppressWarnings(cor(
  env,
  use    = "pairwise.complete.obs",
  method = "spearman"
))

# R squared matrix
r2_mat <- cor_mat^2
```

print correlation matrix
```{r}
colnames(cor_mat) <- env_labels
rownames(cor_mat) <- env_labels
colnames(r2_mat) <- env_labels
rownames(r2_mat) <- env_labels

corrplot(
  cor_mat,
  method = "color",
  type = "upper",
  tl.col = "black",
  tl.cex = 0.8,
  tl.srt = 45,
  addgrid.col = NA,
  col = colorRampPalette(c(
    "#08306B","#2171B5","#6BAED6","#C6DBEF","#F7FBFF",
    "#FFF5F0","#FCBBA1","#FB6A4A","#CB181D","#67000D"
  ))(200),
  
  # add R squared
  addCoef.col = "black",
  number.cex = 0.8,
  number.matrix = round(r2_mat, 2)
)
```


#3. Radar plots of covariate, outcome, and exposure distributions across study centers (Figure S4)

variables
```{r}
age_var <- "u1_basist_age"
bmi_var <- "u1_at_anthro_bmi_eig"
sex_var <- "u1_basist_sex"
popdensity_var <- "u1_gt_pop_d_ink_a_18"
gisd_var <- "u1_gt_gisd_rki_m"

all_vars <- c(
  temp_var,
  pollutant_vars,
  popdensity_var,
  gisd_var,
  age_var,
  #sex_var, #have to add proportion female as a separate varialble so sex is numeric
  bmi_var
)

study_center <- "u1_basist_uort_combined"

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_pop_d_ink_a_18" = "Population\nDensity",
  "copd_incidence" = "COPD",
  "asthma_incidence" = "Asthma",
  "u1_at_anthro_bmi_eig" = "BMI",
  "u1_basist_age" = "Age",
  "prop_female" = "Percent\nFemale"
)
```

```{r}
df <- u2_df_copd


#get all of the variables into one plot
get_incidence <- function(df, outcome) {
  df %>%
    group_by(.data[[study_center]]) %>%
    summarise(incidence = mean(.data[[outcome]] == 1, na.rm = TRUE) * 100)
}

env_summary <- df %>%
  group_by(.data[[study_center]]) %>%
  summarise(
    across(all_vars, ~ mean(.x, na.rm = TRUE)),
    prop_female = mean(u1_basist_sex == 1, na.rm = TRUE) 
  )

copd_inc <- get_incidence(u2_df_copd, "u2_dt_an_lung_1_fu") %>%
  rename(copd_incidence = incidence)

asthma_inc <- get_incidence(u2_df_asthma, "u2_dt_an_lung_2_fu") %>%
  rename(asthma_incidence = incidence)

combined <- env_summary %>%
  left_join(copd_inc, by = study_center) %>%
  left_join(asthma_inc, by = study_center)

# Normalize 0–1
scaled <- combined %>%
  mutate(across(-all_of(study_center), ~ rescale(.x, to = c(0, 1))))

pretty_df <- scaled %>%
  rename_with(~ ifelse(!is.na(labels[.x]), labels[.x], .x))

#create plots
plot_grobs <- list()

centers <- unique(pretty_df[[study_center]])

for (center in centers) {
  
  dat <- pretty_df %>%
    filter(.data[[study_center]] == center) %>%
    select(-all_of(study_center))
  
  radar_dat <- rbind(
    rep(1, ncol(dat)),
    rep(0, ncol(dat)),
    dat
  )
  
  f <- tempfile(fileext = ".png")
  png(f, width = 900, height = 900, res = 140)
  par(cex = 1.35)
  par(mar = c(1.5, 1.5, 4, 1.5))
  
  radarchart(
    radar_dat,
    axistype = 1,
    pcol = "steelblue",
    pfcol = scales::alpha("steelblue", 0.35),
    plwd = 3,
    
    # Grid lines
    cglcol = "grey50",
    cglty = 1,
    cglwd = 1.6,
    vlcex = 1.4,
    caxislabels = rep("", 5),
    title = list(paste(center), cex = 2)
  )
  
  dev.off()
  
  img <- readPNG(f)
  plot_grobs[[center]] <- rasterGrob(img)
}

# Arrange on a compact grid
n <- length(plot_grobs)
ncol <- ceiling(sqrt(n))
nrow <- ceiling(n / ncol)

grid.arrange(
  grobs = plot_grobs,
  ncol = ncol,
  nrow = nrow,
  padding = unit(0, "mm")
)


png("", width = 3500, height = 3500, res = 200)
grid.arrange(
  grobs = plot_grobs,
  ncol = ncol,
  nrow = nrow,
  padding = unit(0, "mm")
)
dev.off()

```


#4. Number of individuals at each examination center (Table S1)

examination center-table
```{r}
# Summary for u1_year
u1_summary <- u1_df %>%
  count(u1_basist_uort) %>%
  mutate(
    pct = round(n / sum(n) * 100, 2),
    u1 = paste0(n, " (", pct, "%)")
  ) %>%
  select(center = u1_basist_uort, u1)

# Summary for u2_year
u2_summary <- u2_df_copd %>%
  count(u2_basist_uort) %>%
  mutate(
    pct = round(n / sum(n) * 100, 2),
    u2 = paste0(n, " (", pct, "%)")
  ) %>%
  select(center = u2_basist_uort, u2)

# Combine into one table with years as rows
exam_center_summary <- full_join(u1_summary, u2_summary, by = "center") %>%
  arrange(center)

write_csv(exam_center_summary, file = file.path("")) #save table
```

examination center-figure
```{r}
plot_center_bar <- function(data, var, var_name) {
  ggplot(data, aes(x = .data[[var]])) +
    geom_bar(fill = "#4C72B0", color = "white") +
    coord_flip() +
    labs(
      title = var_name,
      x = "Study Center",
      y = "Count"
    ) +
    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 6),
      axis.text  = element_text(size = 5)
    )
}

plot <- grid.arrange(
  plot_center_bar(u1_df, "u1_basist_uort", "Baseline Study Center"),
  plot_center_bar(u2_df_copd, "u2_basist_uort", "Follow up 1 Study Center"),
  ncol = 3,
  nrow = 2
)

plot

ggsave(
  filename = file.path(""),plot = plot)
```


#5. Length of follow up time for participants

```{r}
# Time difference in years
time_diff_years <- as.numeric(u2_df_asthma$u2_basist_udat - u2_df_asthma$u1_basist_udat) / 365.25

summary(time_diff_years)

# Histogram
hist(
  time_diff_years,
  breaks = 50,
  main = "Time Between U1 and U2",
  xlab = "Years",
  col = "skyblue",
  border = "white"
)
```


#6. Explore outliners and distribution shapes for continuous variables

function to create a table1 for continuous variables
```{r}
summary_continuous <- function(df, var_label, u1_var) {
  
  u1 <- df[[u1_var]]
  
  # Compute IQR thresholds
  Q1 <- quantile(u1, 0.25, na.rm = TRUE)
  Q3 <- quantile(u1, 0.75, na.rm = TRUE)
  IQR_val <- IQR(u1, na.rm = TRUE)
  
  lower_thr <- Q1 - 1.5 * IQR_val
  upper_thr <- Q3 + 1.5 * IQR_val
  
  # Outliers
  outliers <- u1[u1 < lower_thr | u1 > upper_thr]
  
  tibble::tibble(
    variable = paste0(var_label, ", mean (sd)"),
    distribution = sprintf("%.2f (%.2f)", mean(u1, na.rm = TRUE), sd(u1, na.rm = TRUE)),
    outliers = sprintf("threshold: [%.2f, %.2f], number: %d (%.2f%%) ", lower_thr, upper_thr,
                             length(outliers), 100 * length(outliers) / length(u1))
  )
}
```

function to create histogram figures for continuous variables
```{r}
hist_var <- function(df, var_name, var, bin_width) {
  ggplot(df, aes(x = .data[[var]])) +
    geom_histogram(binwidth = bin_width, fill = "#4C72B0", color = "white") +
    labs(
      title = paste0(var_name),
      x = var_name,
      y = "Count"
    ) +
    theme_minimal(base_size = 8) +   # Increase all base text
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 6),
      axis.text  = element_text(size = 5)
    )
}
```

function to create boxplots for continuous variables
```{r}
boxplot_var <- function(data, var_name, var) {
  ggplot(data, aes(y = .data[[var]])) +
    geom_boxplot(fill = "#4C72B0", color = "black", outlier.color = "red", size = 0.4) +
    labs(
      title = var_name,
      y = var_name,
      x = ""
    ) +
    coord_flip() +
    theme_minimal(base_size = 8) +
    theme(
      plot.title = element_text(size = 10, face = "bold"),
      axis.title = element_text(size = 6),
      axis.text  = element_text(size = 5)
    )
}
```


7.1. cohort data continuous variables
table
```{r}
summary_table <- dplyr::bind_rows(
  summary_continuous(u2_df_copd, "age", "u2_basist_age"),
  summary_continuous(u2_df_copd, "BMI", "u2_at_anthro_bmi"),
  summary_continuous(u2_df_copd, "Physical activity per day", "u2_at_gpaq_ptotalday"),
  summary_continuous(u2_df_copd, "Age of COPD diagnosis", "u2_at_lun_age_copd"),
  summary_continuous(u2_df_copd, "Age of asthma diagnosis", "u2_at_lun_age_asth")
)

summary_table
```

histogram & boxplots
```{r}
plot <- grid.arrange(
  hist_var(u2_df_copd, "age", "u2_basist_age", 5),
  hist_var(u2_df_copd, "BMI", "u2_at_anthro_bmi", 3),
  hist_var(u2_df_copd, "Physical activity per day", "u2_at_gpaq_ptotalday", 100),
  hist_var(u2_df_copd, "Age of COPD diagnosis", "u2_at_lun_age_copd", 3),
  hist_var(u2_df_copd, "Age of asthma diagnosis", "u2_at_lun_age_asth", 3),
  ncol = 3,
  nrow = 3
)

plot

plot <- grid.arrange(
  boxplot_var(u2_df_copd, "age", "u2_basist_age"),
  boxplot_var(u2_df_copd, "BMI", "u2_at_anthro_bmi"),
  boxplot_var(u2_df_copd, "Physical activity per day", "u2_at_gpaq_ptotalday"),
  boxplot_var(u2_df_copd, "Age of COPD diagnosis", "u2_at_lun_age_copd"),
  boxplot_var(u2_df_copd, "Age of asthma diagnosis", "u2_at_lun_age_asth"),
  ncol = 3,
  nrow = 5
)

plot

```

7.2. environmental variables
```{r}
plot <- grid.arrange(
  hist_var(u1_df, "Avg Temp",              "u1_gt_t_mean_hmg_g",        0.25),
  hist_var(u1_df, "3-year Avg Temp",       "u1_gt_t_mean_hmg_g_3year",  0.25),
  hist_var(u1_df, "Fall Avg Temp",         "u1_gt_t_fal_mean_hmg_g",    0.25),
  hist_var(u1_df, "Winter Avg Temp",       "u1_gt_t_win_mean_hmg_g",    0.25),
  hist_var(u1_df, "Spring Avg Temp",       "u1_gt_t_spr_mean_hmg_g",    0.25),
  hist_var(u1_df, "Summer Avg Temp",       "u1_gt_t_smr_mean_hmg_g",    0.25),
  ncol = 3,
  nrow = 3
)
plot

plot <- grid.arrange(
  hist_var(u1_df, "Temp SD",               "u1_gt_t_sd_hmg_g",          0.25),
  hist_var(u1_df, "Temp Range",            "u1_gt_t_rng_mean_hmg_g",    0.25),
  hist_var(u1_df, "Temp Max",              "u1_gt_t_max_hmg_g",         0.25),
  hist_var(u1_df, "Temp Min",              "u1_gt_t_min_hmg_g",         0.25),
  hist_var(u1_df, "Heat Days",             "u1_gt_heatd_hmg_g",         1),
  hist_var(u1_df, "Frost Days",            "u1_gt_frostd_hmg_g",        2),
  ncol = 3,
  nrow = 3
)
plot

plot <- grid.arrange(
  hist_var(u1_df, "PM2.5",                 "u1_gt_pm25_exp_g",          1),
  hist_var(u1_df, "PM10",                  "u1_gt_pm10_exp_g",          1),
  hist_var(u1_df, "Ozone",                 "u1_gt_o3_exp_g",            3),
  hist_var(u1_df, "NO2",                   "u1_gt_no2_exp_g",           3),
  hist_var(u1_df, "Global Radiation",      "u1_gt_rad_dwd_g",           10),
  hist_var(u1_df, "Avg Humidity",          "u1_gt_rh_mean_hmg_g",       1),
  hist_var(u1_df, "Humidity SD",           "u1_gt_rh_sd_hmg_g",         0.5),
  ncol = 3,
  nrow = 3
)
plot
```

```{r}
plot <- grid.arrange(
  boxplot_var(u1_df, "Avg Temp",              "u1_gt_t_mean_hmg_g"),
  boxplot_var(u1_df, "3-year Avg Temp",       "u1_gt_t_mean_hmg_g_3year"),
  boxplot_var(u1_df, "Fall Avg Temp",       "u1_gt_t_fal_mean_hmg_g"),
  boxplot_var(u1_df, "Winter Avg Temp",       "u1_gt_t_win_mean_hmg_g"),
  boxplot_var(u1_df, "Spring Avg Temp",       "u1_gt_t_spr_mean_hmg_g"),
  boxplot_var(u1_df, "Summer Avg Temp",       "u1_gt_t_smr_mean_hmg_g"),
  boxplot_var(u1_df, "Temp SD",               "u1_gt_t_sd_hmg_g"),
  boxplot_var(u1_df, "Temp Range",            "u1_gt_t_rng_mean_hmg_g"),
  boxplot_var(u1_df, "Temp Max",              "u1_gt_t_max_hmg_g"),
  boxplot_var(u1_df, "Temp Min",              "u1_gt_t_min_hmg_g"),
  boxplot_var(u1_df, "Heat Days",             "u1_gt_heatd_hmg_g"),
  boxplot_var(u1_df, "Frost Days",            "u1_gt_frostd_hmg_g"),
  ncol = 3,
  nrow = 4
)
plot

plot <- grid.arrange(
  boxplot_var(u1_df, "PM2.5",                 "u1_gt_pm25_exp_g"),
  boxplot_var(u1_df, "PM10",                  "u1_gt_pm10_exp_g"),
  boxplot_var(u1_df, "Ozone",                 "u1_gt_o3_exp_g"),
  boxplot_var(u1_df, "NO2",                   "u1_gt_no2_exp_g"),
  boxplot_var(u1_df, "Global Radiation",      "u1_gt_rad_dwd_g"),
  boxplot_var(u1_df, "Avg Humidity",          "u1_gt_rh_mean_hmg_g"),
  boxplot_var(u1_df, "Humidity SD",           "u1_gt_rh_sd_hmg_g"),
  ncol = 3,
  nrow = 3
)
plot
```


