
This file is used to examine missingness paterns and impute missing data.

```{r}
library(randomForest)
library(missForest)
library(doParallel)
library(foreach)
library(dplyr)
library(purrr)
library(stringr)
```


#1. Load data
```{r}
u1_df        <- readRDS("") #baseline df
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline
```


#2. Check missing at random for variables we want to impute

get the number of missing for each variable
```{r}
pct_missing <- function(df, vars) {
  round(
    sapply(vars, function(v) {
      if (!v %in% names(df)) {
        return(NA_real_)   # variable not present
      }
      mean(is.na(df[[v]])) * 100
    }),
    2
  )
}

vars <- c("u1_at_ses_mig_status","u1_at_smok_stat_noint","u1_at_ses_isced11_level","u1_at_ses_inc","u1_d_bb1","u1_gt_inc_hh_ink_d","u1_gt_gisd_rki_m","u1_gt_dgu_eu_m","u1_gt_pop_d_ink_a_18","u1_gt_ndvi_300_exp_g_15","u1_at_anthro_bmi_eig","u2_at_anthro_bmi_eig","u1_at_gpaq_ptotal_quartile","u2_at_gpaq_ptotal_quartile","u2_dt_an_lung_1_fu","u2_dt_an_lung_2_fu","u1_d_an_lung_2","u1_d_an_lung_1")

missing_table <- data.frame(
  variable = vars,
  u1_df = pct_missing(u1_df, vars),
  u2_df_asthma   = pct_missing(u2_df_asthma, vars),
  u2_df_copd     = pct_missing(u2_df_copd, vars),
  row.names = NULL
)
```

heatmap of missingness
- some correlated missings between income, occupational exposure, smoking, and education
```{r}
#Plot missing pattern: 
pattern <- u1_df[, c("u1_at_ses_mig_status","u1_at_smok_stat_noint","u1_at_ses_isced11_level","u1_at_ses_inc","u1_d_bb1","u1_gt_inc_hh_ink_d","u1_gt_gisd_rki_m","u1_gt_dgu_eu_m","u1_gt_pop_d_ink_a_18","u1_gt_ndvi_300_exp_g_15","u1_at_anthro_bmi_eig","u1_at_gpaq_ptotal_quartile", "u1_basist_age", "u1_basist_sex", "u1_basist_uort", "u1_d_an_lung_2", "u1_d_an_lung_1")]

colnames(pattern) <- c("migration", "smoking", "education", "income", "occupational exposure", "household income", "GISD", "urbanization", "pop density", "NDVI", "BMI", "physical activity", "age", "sex","study center", "asthma", "COPD")

#create "heatmap" with missings
ggplot2::ggsave(
  "",
  plot = plot_pattern(pattern, rotate = TRUE),
  width = 14,
  height = 10,
  dpi = 150
)
```

helper methods to use visualizations to look for missing at random
```{r}
u1_missingness_plots <- function(data, var_name, lim) {
  
  # Create missingness indicator
  data$missing_flag <- ifelse(is.na(data[[var_name]]), "Missing", "Not missing")
  
  # Helper function to build each plot
  plot_missing <- function(xvar, title, xlab) {
    ggplot(data, aes(x = .data[[xvar]], fill = missing_flag)) +
      geom_bar(position = "fill") +
      scale_y_continuous(labels = percent) +
      coord_cartesian(ylim = c(lim, 1)) + 
      labs(
        title = NULL,
        x = xlab,
        y = NULL
      ) +
      theme(
        plot.title = element_text(size = 6),
        axis.title = element_text(size = 6),
        axis.text.x  = element_text(size = 6),
        axis.text.y  = element_text(size = 6),
        legend.position = "none"
      )
  }
  
  # Return all 5 plots
  list(
    age   = plot_missing("u1_basist_age",  "Missingness", "Age"),
    sex   = plot_missing("u1_basist_sex",  "Missingness", "Sex"),
    uort  = plot_missing("u1_basist_uort", "Missingness", "Study Center"),
    lung2 = plot_missing("u1_d_an_lung_2", "Missingness", "Asthma"),
    lung1 = plot_missing("u1_d_an_lung_1", "Missingness", "COPD")
  )
}

u2_missingness_plots <- function(data, var_name, lim) {
  
  # Create missingness indicator
  data$missing_flag <- ifelse(is.na(data[[var_name]]), "Missing", "Not missing")
  
  # Helper function to build each plot
  plot_missing <- function(xvar, title, xlab) {
    ggplot(data, aes(x = .data[[xvar]], fill = missing_flag)) +
      geom_bar(position = "fill") +
      scale_y_continuous(labels = percent) +
      coord_cartesian(ylim = c(lim, 1)) + 
      labs(
        title = NULL,
        x = xlab,
        y = NULL
      ) +
      theme(
        plot.title = element_text(size = 6),
        axis.title = element_text(size = 6),
        axis.text.x  = element_text(size = 6),
        axis.text.y  = element_text(size = 6),
        legend.position = "none"
      )
  }
  
  # Return all 5 plots
  list(
    age   = plot_missing("u1_basist_age",  "Missingness", "Age"),
    sex   = plot_missing("u1_basist_sex",  "Missingness", "Sex"),
    uort  = plot_missing("u1_basist_uort", "Missingness", "Study Center"),
    lung2 = plot_missing("u2_dt_an_lung_2_fu", "Missingness", "Asthma"),
    lung1 = plot_missing("u2_dt_an_lung_1_fu", "Missingness", "COPD")
  )
}
```

looking for missing at random for bmi & physical activity (baseline and follow up)
- increased missing at older ages
```{r}
plots <- u1_missingness_plots(u1_df, "u1_at_anthro_bmi_eig", 0.75)
(plots$age | plots$sex | plots$uort | plots$lung2 | plots$lung1) /
plot_spacer() / 
plot_spacer()

plots <- u2_missingness_plots(u2_df_asthma, "u2_at_anthro_bmi_eig", 0.75)
(plots$age | plots$sex | plots$uort | plots$lung2 | plots$lung1) /
plot_spacer() / 
plot_spacer()

plots <- u1_missingness_plots(u1_df, "u1_at_gpaq_ptotal_quartile", 0.65)
(plots$age | plots$sex | plots$uort | plots$lung2 | plots$lung1) /
plot_spacer() / 
plot_spacer()

plots <- u2_missingness_plots(u2_df_asthma, "u2_at_gpaq_ptotal_quartile", 0.65)
(plots$age | plots$sex | plots$uort | plots$lung2 | plots$lung1) /
plot_spacer() / 
plot_spacer()

```

looking for missing at random for migration, smoking, education, income, occupational exposure, neighborhood income, ndvi, population density, degree of urbanization (only baseline)
- also some evidence of increased missing at older ages
- 100% missing for Leipzig in 2015 -> use 2018 for all
```{r}
# 1
plots_mig <- u1_missingness_plots(u1_df, "u1_at_ses_mig_status", 0.9)
(plots_mig$age | plots_mig$sex | plots_mig$uort | plots_mig$lung2 | plots_mig$lung1) /
  plot_spacer() / plot_spacer()

# 2
plots_smok <- u1_missingness_plots(u1_df, "u1_at_smok_stat_noint", 0.7)
(plots_smok$age | plots_smok$sex | plots_smok$uort | plots_smok$lung2 | plots_smok$lung1) /
  plot_spacer() / plot_spacer()

# 3
plots_isced <- u1_missingness_plots(u1_df, "u1_at_ses_isced11_level", 0.9)
(plots_isced$age | plots_isced$sex | plots_isced$uort | plots_isced$lung2 | plots_isced$lung1) /
  plot_spacer() / plot_spacer()

# 4
plots_inc <- u1_missingness_plots(u1_df, "u1_at_ses_inc", 0.75)
(plots_inc$age | plots_inc$sex | plots_inc$uort | plots_inc$lung2 | plots_inc$lung1) /
  plot_spacer() / plot_spacer()

# 5
plots_bb1 <- u1_missingness_plots(u1_df, "u1_d_bb1", 0.15)
(plots_bb1$age | plots_bb1$sex | plots_bb1$uort | plots_bb1$lung2 | plots_bb1$lung1) /
  plot_spacer() / plot_spacer()

# 6
plots_hhinc <- u1_missingness_plots(u1_df, "u1_gt_inc_hh_ink_d", 0.75)
(plots_hhinc$age | plots_hhinc$sex | plots_hhinc$uort | plots_hhinc$lung2 | plots_hhinc$lung1) /
  plot_spacer() / plot_spacer()

# 7
plots_gisd <- u1_missingness_plots(u1_df, "u1_gt_gisd_rki_m", 0.75)
(plots_gisd$age | plots_gisd$sex | plots_gisd$uort | plots_gisd$lung2 | plots_gisd$lung1) /
  plot_spacer() / plot_spacer()

# 8
plots_dgu <- u1_missingness_plots(u1_df, "u1_gt_dgu_eu_m", 0.75)
(plots_dgu$age | plots_dgu$sex | plots_dgu$uort | plots_dgu$lung2 | plots_dgu$lung1) /
  plot_spacer() / plot_spacer()

# 9
plots_pop <- u1_missingness_plots(u1_df, "u1_gt_pop_d_ink_a_18", 0.25)
(plots_pop$age | plots_pop$sex | plots_pop$uort | plots_pop$lung2 | plots_pop$lung1) /
  plot_spacer() / plot_spacer()

# 10
plots_ndvi <- u1_missingness_plots(u1_df, "u1_gt_ndvi_300_exp_g_15", 0.75)
(plots_ndvi$age | plots_ndvi$sex | plots_ndvi$uort | plots_ndvi$lung2 | plots_ndvi$lung1) /
  plot_spacer() / plot_spacer()
```




#3. Prepare dataset for imputation

exclude variables related to the outcomes of interest--we do not want to impute these
- Dad asthma history
- Mom asthma history
- Residential history
- COPD medication
- Asthma medication
- Age of asthma diagnosis
- Age of COPD diagnosis

exlcude other variables that will slow down the algorithm and not important predictors
- id
- exam date
- income categories (we can calculate these after if we impute income)
- physical activity (we only want the transformed categorical variable)
```{r}
exclude_vars_u1 <- c("ID", "u1_basist_udat", "u1_basist_year", "u1_dt_fa_v3_ab", "u1_dt_fa_m3_ab", "u1_d_umw2_j", "u1_at_lun_age_asth", "u1_at_lun_age_copd", "u1_dt_an_lungasth2", "u1_dt_an_lungcopd2", "u1_dt_lg2_2", "u1t_l388lq_medic4", "u1_at_ses_inc_categories", "u1_at_ses_inc_poverty", "u1_at_ses_inc_median", "u1_at_ses_isced11_3level", "u1_d_an_lung_2", "u1_d_an_lung_1")
exclude_vars_u2 <- c("u2_basist_udat", "u2_basist_year", "u2_dt_fa_v3_ab", "u2_dt_fa_m3_ab",  "u2_d_umw2_j", "u2_at_lun_age_asth","u2_at_lun_age_copd","u2_dt_an_lungasth2_fu", "u2_dt_an_lung_2_fu", "u2_dt_an_lung_1_fu")

u2_df_asthma_imputation <- u2_df_asthma %>%
  select(-all_of(c(exclude_vars_u1, exclude_vars_u2)))

u2_df_copd_imputation <- u2_df_copd %>%
  select(-all_of(c(exclude_vars_u1, exclude_vars_u2)))
```

confirm that the exclusion works corrrectly
```{r}
missing_table_df <- data.frame(
  variable = names(u1_df_imputation),
  missing_count = sapply(u1_df_imputation, function(x) sum(is.na(x)))
)

missing_table_df

missing_table_df2 <- data.frame(
  variable = names(u2_df_asthma_imputation),
  missing_count = sapply(u2_df_asthma_imputation, function(x) sum(is.na(x)))
)

missing_table_df2
```

check that all categorical variables are factors
```{r}
u2_df_asthma_imputation %>%
  summarise(across(everything(), ~ class(.x))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "datatype")
```


#4. Imputation of follow-up dataset: Asthma

```{r}
no_cores <- detectCores() - 2     #leave 2 cores left  
cl <- makeCluster(no_cores)  
registerDoParallel(cl)

set.seed(809)   # reproducibility

mf_result_100_fu_asthma <- missForest(
  u2_df_asthma_imputation,
  maxiter = 10,        # default iterations
  ntree = 100,         # default RF size
  variablewise = TRUE, # gives per-variable error
  verbose = TRUE,       # print progress
  parallelize = "forests"
)

# Extract the imputed dataset
imputed_100_fu_asthma <- mf_result_100_fu_asthma$ximp

# Inspect OOB error
mf_result_100_fu_asthma$OOBerror
```

checking accuracy of the imputed variables using OOB error
```{r}
imp100_result_asthma <- data.frame(
  variable = names(u2_df_asthma_imputation),
  missing = colSums(is.na(u2_df_asthma_imputation)),
  OOB_error = mf_result_100_fu_asthma$OOBerror
)[order(-mf_result_100_fu_asthma$OOBerror), ]

imp100_result_asthma$OOB_error <- round(imp100_result_asthma$OOB_error, 3)
```

add the excluded columns back in and save
```{r}
# Extract excluded variables BEFORE imputation
excluded_part <- u2_df_asthma %>% select(any_of(c(exclude_vars_u1, exclude_vars_u2)))

# Add excluded variables back AFTER imputation
u2_df_asthma_final <- bind_cols(
  imputed_100_fu_asthma,
  excluded_part
)
```

```{r}
saveRDS(u2_df_asthma_final,file = "")
```

#5 Imputation of follow-up dataset: COPD

```{r}
no_cores <- detectCores() - 2     #leave 2 cores left  
cl <- makeCluster(no_cores)  
registerDoParallel(cl)

set.seed(809)   # reproducibility

mf_result_100_fu_copd<- missForest(
  u2_df_copd_imputation,
  maxiter = 10,        # default iterations
  ntree = 100,         # default RF size
  variablewise = TRUE, # gives per-variable error
  verbose = TRUE,       # print progress
  parallelize = "forests"
)

# Extract the imputed dataset
imputed_100_fu_copd <- mf_result_100_fu_copd$ximp

# Inspect OOB error
mf_result_100_fu_copd$OOBerror
```

checking accuracy of the imputed variables using OOB
```{r}
imp100_result_copd <- data.frame(
  variable = names(u2_df_copd_imputation),
  missing = colSums(is.na(u2_df_copd_imputation)),
  OOB_error = mf_result_100_fu_copd$OOBerror
)[order(-mf_result_100_fu_copd$OOBerror), ]

imp100_result_copd$OOB_error <- round(imp100_result_copd$OOB_error, 3)
```

add the excluded columns back in and save
```{r}
# Extract excluded variables BEFORE imputation
excluded_part <- u2_df_copd %>% select(any_of(c(exclude_vars_u1, exclude_vars_u2)))

# Add excluded variables back AFTER imputation
u2_df_copd_final <- bind_cols(
  imputed_100_fu_copd,
  excluded_part
)
```

```{r}
saveRDS(u2_df_copd_final,file="")
```


#6 Testing with more random forests
- 500 forests runtime about 3 hours
- did not increase the accuracy meaningfully
```{r}
X2 <- u1_df_imputation
```

```{r}
no_cores <- detectCores() - 2     #leave 2 cores left  
cl <- makeCluster(no_cores)  
registerDoParallel(cl)

set.seed(809)   # reproducibility

mf_result <- missForest(
  X2,
  maxiter = 10,        # default iterations
  ntree = 500,         # default RF size
  variablewise = TRUE, # gives per-variable error
  verbose = TRUE,       # print progress
  parallelize = "forests"
)

# Extract the imputed dataset
X2_imputed <- mf_result$ximp

# Inspect OOB error
mf_result$OOBerror

# save
write.csv(X2_imputed, "", row.names = FALSE)
```

compare results
```{r}
# Create tables (do NOT print yet)
desc_cc  <- CreateTableOne(vars = names(X), data = X)
desc_imp100 <- CreateTableOne(vars = names(X_imputed), data = X_imputed)
desc_imp500 <- CreateTableOne(vars = names(X2_imputed), data = X2_imputed)

# Convert to data frames
df_cc  <- print(desc_cc,  printToggle = FALSE, noSpaces = TRUE)
df_imp100 <- print(desc_imp100, printToggle = FALSE, noSpaces = TRUE)
df_imp500 <- print(desc_imp500, printToggle = FALSE, noSpaces = TRUE)

# Combine into one data frame
desc <- data.frame(
  Variable = rownames(df_cc),
  Complete_case = df_cc[, 1],
  Imputed_DS_100 = df_imp[, 1],
  Imputed_DS_500 = df_imp[, 1],
  row.names = NULL
)

desc
```

checking accuracy of the imputed variables
```{r}
imp500_result <- data.frame(
  variable = names(X),
  missing = colSums(is.na(X)),
  OOB_error = mf_result$OOBerror
)[order(-mf_result$OOBerror), ]

imp500_result$OOB_error <- round(imp500_result$OOB_error, 3)
```




