
This file contains all the code to run the main analyses and models presented in the paper

```{r}
library(dplyr)
library(mgcv)
library(purrr)
library(tibble)
library(patchwork)
library(ggplot2)
library(future)
library(future.apply)
library(gWQS)
library(ggsci)
library(cowplot)
library(scales)
```


# Load data
```{r}
#main datasets
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline
```


# Setting up variables

setting up to run code in parallel
```{r}
plan(
  multisession,
  workers = max(1, parallel::detectCores() - 1)
)

options(future.globals.maxSize = 8 * 1024^3)
```

setting up scenarios--one for COPD, one for asthma
```{r}
scenarios <- list(
  copd_incidence = list(df = u2_df_copd, outcome = "u2_dt_an_lung_1_fu", title = "COPD Incidence"),
  asthma_incidence = list(df = u2_df_asthma, outcome = "u2_dt_an_lung_2_fu", title = "Asthma Incidence")
)
```

setting up variables
```{r}
temp_var <- "u1_gt_t_mean_hmg_g"
winter_var <- "u1_gt_t_win_mean_hmg_g"
temp_vars <- c("u1_gt_t_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g")
seasons_order <- c("u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g")
humidity_var <- "u1_gt_rh_mean_hmg_g"

pm_vars <- c("u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g")
no2_var <-"u1_gt_no2_exp_g"
pollutant_vars <- c(pm_vars, no2_var)

all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)

study_center_var <-"u1_basist_uort_combined"
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)

titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)
```




#1. GAMMS with non-linear terms for environmental data
- For Figure 1 and Figure S5

variable set up -- temp & seasons
```{r}
all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)
```

variable set up -- all temperature seasons
```{r}
all_vars <- c(temp_var, seasons_order)
```

function -- final version
```{r}
fit_gamm <- function(
    y,  data,
    x,
    label,
    title,
    covars,
    study_center_random_int = FALSE
) {
  
  #ensure study center is factor (for random intercept)
  if (study_center_random_int) {
    data[[study_center_var]] <- factor(data[[study_center_var]])
  }
  
  #build formula
  covar_string <- paste(covars, collapse = " + ")
  
  re_term <- if (study_center_random_int) {
    paste0(" + s(", study_center_var, ", bs = 're')")
  } else {
    ""
  }
  
  form <- as.formula(
    paste0(
      y,
      " ~ s(", x, ", k=4) + ",
      covar_string,
      re_term
    )
  )
  
  #fit GAM
  fit <- mgcv::gam(
    formula = form,
    data    = data,
    family  = binomial(link = "logit"),
    method  = "REML"
  )
  
  #prediction grid 
  xseq <- seq(
    min(data[[x]], na.rm = TRUE),
    max(data[[x]], na.rm = TRUE),
    length.out = 200
  )
  
  newd <- data.frame(exposure_value = xseq)
  names(newd)[1] <- x
  
  #set covariates to median / mode
  for (v in covars) {
    if (is.numeric(data[[v]])) {
      newd[[v]] <- median(data[[v]], na.rm = TRUE)
    } else {
      mode_val <- names(sort(table(data[[v]]), decreasing = TRUE))[1]
      newd[[v]] <- factor(mode_val, levels = levels(data[[v]]))
    }
  }
  
  if (study_center_random_int) {
    newd[[study_center_var]] <- levels(data[[study_center_var]])[1]
  }
  
  #predict on log-odds scale
  pred <- predict(
    fit,
    newdata = newd,
    se.fit  = TRUE,
    type    = "link",
    exclude = if (study_center_random_int)
      paste0("s(", study_center_var, ")")
    else
      NULL
  )
  
  #reference = 1st percentile
  ref_val <- stats::quantile(data[[x]], 0.01, na.rm = TRUE)
  ref_idx <- which.min(abs(xseq - ref_val))
  
  #return odds ratios
  return(list(
    preds = tibble::tibble(
      exposure_value = xseq,
      OR      = exp(pred$fit - pred$fit[ref_idx]),
      OR_low  = exp((pred$fit - 1.96 * pred$se.fit) - pred$fit[ref_idx]),
      OR_high = exp((pred$fit + 1.96 * pred$se.fit) - pred$fit[ref_idx]),
      exposure = x,
      label    = label,
      title = title
    ),
    fit = fit
  ))
}
```

function to loop through all environmental variables
- 1st-99th percentiles
- standardizing y-axis
- with histogram
```{r}
gamm_singleexposure_outputs <- lapply(
  scenarios,
  function(s) {
    
    plotdata_list <- future.apply::future_lapply(
      all_vars,
      function(x) {
        fit_gamm(
          data  = s$df,
          y     = s$outcome,
          x     = x,
          label = if (x %in% names(labels)) labels[[x]] else x,
          title = if (x %in% names(titles)) titles[[x]] else x,
          covars = covars,
          study_center_random_int = TRUE
        )
      },
      future.seed = TRUE
    )
    
    names(plotdata_list) <- sapply(
      all_vars,
      function(x) {
        if (x %in% names(labels)) labels[[x]] else x
      }
    )
    
    plots <- lapply(
      seq_along(plotdata_list),
      function(i) {
        
        df <- plotdata_list[[i]]$preds
        
        lab   <- df$label[1]
        title <- df$title[1]
        xvar  <- df$exposure[1]
        
        # percentiles from original data
        p1  <- quantile(s$df[[xvar]], 0.01, na.rm = TRUE)
        p99 <- quantile(s$df[[xvar]], 0.99, na.rm = TRUE)
        p5  <- quantile(s$df[[xvar]], 0.05, na.rm = TRUE)
        p95 <- quantile(s$df[[xvar]], 0.95, na.rm = TRUE)
        
        # restrict GAMM curve
        df <- df %>%
          dplyr::filter(
            exposure_value >= p1,
            exposure_value <= p99
          )
        
        # histogram data
        hist_df <- s$df %>%
          dplyr::filter(
            .data[[xvar]] >= p1,
            .data[[xvar]] <= p99
          )
        
        # GAMM plot (top panel)
        p_gamm <- ggplot(
          df,
          aes(x = exposure_value, y = OR)
        ) +
          geom_ribbon(
            aes(
              ymin = OR_low,
              ymax = OR_high
            ),
            fill = "grey50",
            alpha = 0.3
          ) +
          geom_line(
            linewidth = 1.1,
            color = "black"
          ) +
          geom_hline(
            yintercept = 1,
            #linetype = "dashed",
            color = "gray30",
            linewidth = 0.5
          ) +
          geom_vline(
            xintercept = c(p5, p95),
            linetype = "dashed",
            color = "gray50",
            linewidth = 0.5
          ) +
          coord_cartesian(
            xlim = c(p1, p99)
          ) +
          scale_y_log10(
            limits = c(0.45, 2.8)
          ) +
          labs(
            title = title,
            x = NULL,
            y = "Odds Ratio"
          ) +
          #theme_minimal(base_size = 9) +   #for all seasons plot
          theme_minimal(base_size = 12) +
          theme(
            #plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            plot.margin = margin(5, 5, 0, 5),
            axis.line.y = element_line(color = "gray50", linewidth = 0.5),
          )
        
        # Histogram (bottom panel)
        p_hist <- ggplot(
          hist_df,
          aes(x = .data[[xvar]])
        ) +
          geom_histogram(
            bins = 30,
            fill = "grey75",
            colour = "white",
            linewidth = 0.2
          ) +
          geom_vline(
            xintercept = c(p5, p95),
            linetype = "dashed",
            color = "gray50",
            linewidth = 0.5
          ) +
          coord_cartesian(
            xlim = c(p1, p99)
          ) +
          labs(
            x = lab,
            y = "Count"
          ) +
          #theme_minimal(base_size = 6.5) +   #for all seasons plot
          theme_minimal(base_size = 10) +
          theme(
            plot.margin = margin(0, 5, 5, 5),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.line.y = element_line(color = "gray50", linewidth = 0.5),
            axis.line.x = element_line(color = "gray50", linewidth = 0.5),
            #axis.title.y = element_text(size = 7),
            #axis.text.y = element_text(size = 6)
            axis.text.y  = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank()
          )
        
        # Combine panels
        p_gamm /
          p_hist +
          patchwork::plot_layout(
            heights = c(9, 1)
          )
      }
    )
    
    list(
      plotdata_list = plotdata_list,
      plots = plots
    )
  }
)
```


plot
```{r}
scenario_boards <- lapply(names(gamm_singleexposure_outputs), function(scn_name) {
  scn   <- gamm_singleexposure_outputs[[scn_name]]
  plots <- scn$plots
  
  board <- wrap_plots(
    plots,
    ncol = 5,
    nrow = 3
  )
  
  board
})

names(scenario_boards) <- names(gamm_singleexposure_outputs)
```

```{r}
scenario_boards$asthma_incidence
scenario_boards$copd_incidence
```


save Figure 1
```{r}
ggsave(
  "",
  plot = scenario_boards$asthma_incidence,
  width = 8,
  height = 6,
  dpi = 1200
)

ggsave(
  "",
  plot = scenario_boards$copd_incidence,
  width = 8,
  height = 6,
  dpi = 1200
)
```

save all seasons plot (Figure S5)
```{r}
ggsave(
  "",
  plot = scenario_boards$asthma_incidence,
  width = 13,
  height = 6,
  dpi = 1200
)

ggsave(
  "",
  plot = scenario_boards$copd_incidence,
  width = 13,
  height = 6,
  dpi = 1200
)
```



#2 adjusted logistic regression with linear term to get ORs
- For Table 2

main function - scale factor is IQR
```{r}
fit_logistic_adjusted <- function(
    df,
    y,
    x,
    covars,
    study_center_random_int = FALSE
) {
  
  # choose scale factor
  # SD
  #sd_x <- stats::sd(df[[x]], na.rm = TRUE)
  #scaled_name <- paste0(x, "_sd")
  #df[[scaled_name]] <- df[[x]] / sd_x
  
  #IQR
  iqr_x <- stats::IQR(df[[x]], na.rm = TRUE)
  scaled_name <- paste0(x, "_iqr")
  df[[scaled_name]] <- df[[x]] / iqr_x
  
  fixed_terms <- c(scaled_name, covars)
  
  
  if (study_center_random_int) {
    fixed_part <- paste(fixed_terms, collapse = " + ")
    rand_part  <- paste0("(1 | ", study_center_var, ")")
    
    form <- as.formula(
      paste(y, "~", fixed_part, "+", rand_part)
    )
    fit <- lme4::glmer(
      form,
      data    = df,
      family  = binomial(),
      control = lme4::glmerControl(
        optimizer   = "bobyqa",
        calc.derivs = FALSE
      )
    )
    
    summ <- summary(fit)$coefficients
    
  } else {
    form <- reformulate(fixed_terms, response = y)
    
    fit <- glm(
      form,
      data   = df,
      family = binomial()
    )
    
    summ <- summary(fit)$coefficients
  }
  
  beta <- summ[scaled_name, "Estimate"]
  se   <- summ[scaled_name, "Std. Error"]
  z    <- qnorm(0.975)
  
  CI_low  <- beta - z * se
  CI_high <- beta + z * se
  p_value <- summ[scaled_name, "Pr(>|z|)"]
  
  OR_val <- exp(beta)
  CI_l   <- exp(CI_low)
  CI_h   <- exp(CI_high)
  
  tibble(
    exposure = x,
    OR       = OR_val,
    CI_low   = CI_l,
    CI_high  = CI_h,
    OR_CI    = sprintf("%.3f (%.3f, %.3f)", OR_val, CI_l, CI_h),
    p_value  = round(p_value, 3)
  )
}
```

new labels
```{r}
titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual\nTemperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer\nTemperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall\nTemperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring\nTemperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)"
)

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual\nTemperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer\nTemperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall\nTemperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring\nTemperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)"
)
```

loop with only final covariates to create Table 2
```{r}
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")
```

loop
```{r}
adjusted_ors2 <- lapply(
  scenarios,
  function(s) {
    
    # Parallelize the model fits (main bottleneck)
    or_table_scaled <- future.apply::future_lapply(
      all_vars,
      function(x) {
        fit_logistic_adjusted(
          df     = s$df,
          y      = s$outcome,
          x      = x,
          covars = covars,
          study_center_random_int = TRUE
        )
      }
    ) |>
      dplyr::bind_rows()
    
    # clean up tables for plotting
    plot_df <- or_table_scaled |>
      dplyr::mutate(
        pretty = factor(labels[exposure], levels = labels[exposure])
      )
    
    #plot
    plt <- ggplot(plot_df, aes(x = pretty, y = OR)) +
      geom_point(size = 3, color = "black") +
      geom_errorbar(
        aes(ymin = CI_low, ymax = CI_high),
        width = 0.12,
        color = "black"
      ) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
      scale_y_log10() +
      labs(
        title = s$title,
        x = "",
        y = "Odds Ratio"
      ) +
      theme_minimal(base_size = 16) +
      theme(
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 13),
        plot.title  = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.line.y = element_line(color = "gray50", linewidth = 0.5),
        axis.line.x = element_line(color = "gray50", linewidth = 0.5)
      )
    
    list(
      or_table = or_table_scaled,
      plot_df  = plot_df,
      plot     = plt
    )
  }
)
```

print
```{r}
adjusted_ors2$asthma_incidence$plot
adjusted_ors2$copd_incidence$plot

asthma_table <- adjusted_ors2$asthma_incidence$or_table
copd_table <- adjusted_ors2$copd_incidence$or_table
```

save
```{r}
write.csv(asthma_table, "asthma_table.csv", row.names = FALSE)
write.csv(copd_table,   "copd_table.csv",   row.names = FALSE)
```


#3 grouped weighted quantile sum regression
- For Figure 2 and Table S3

variables
- note that the gwas does not allow a random intercept so we include the study center as a covariate
```{r}
covars_gwas <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18",
  "u1_basist_uort_combined")
```


#3.1 COPD (winter temperature)
```{r}
wqs_vars <- c(
  "u1_gt_t_win_mean_hmg_g",
  pm_vars,
  no2_var
)
```

```{r}
df           = u2_df_copd
outcome      = "u2_dt_an_lung_1_fu"

df$outcome_bin <- as.integer(df[[outcome]] == 1)

# Keep only needed columns
wqs_df <- df[, c("outcome_bin", wqs_vars, covars_gwas)]
```

run model
```{r}
fit_wqs_copd <- gwqs(
  formula = outcome_bin ~ wqs + 
    u1_basist_sex +
    u1_basist_age +
    u1_at_ses_mig_status +
    u1_at_smok_stat_noint +
    u1_at_ses_isced11_3level +
    u1_gt_gisd_rki_m +
    u1_gt_pop_d_ink_a_18 +
    u1_basist_uort_combined,
  mix_name = wqs_vars,
  data = wqs_df,
  q = 10,
  validation = 0.6,
  b = 100,
  family = "binomial",
  seed = 809,
  tables = TRUE
)
```


get the weights for each environmental variable
```{r}
fit_wqs_copd$final_weights
```

convert results to IQR scale
```{r}
# coefficient table
coef_tab <- summary(fit_wqs_copd$fit)$coefficients

# extract WQS term
beta <- coef_tab["wqs", "Estimate"]
se   <- coef_tab["wqs", "Std. Error"]

# IQR of observed WQS index
iqr_wqs <- IQR(fit_wqs_copd$fit$model$wqs, na.rm = TRUE)

# inference
z_val   <- beta / se
p_value <- 2 * (1 - pnorm(abs(z_val)))
z_crit  <- qnorm(0.975)

# OR per IQR increase in the WQS index
mixture_effect_copd <- tibble::tibble(
  exposure = "mixture",
  IQR_WQS  = iqr_wqs,
  OR       = exp(beta * iqr_wqs),
  CI_low   = exp((beta - z_crit * se) * iqr_wqs),
  CI_high  = exp((beta + z_crit * se) * iqr_wqs),
  p_value  = p_value,
  OR_CI    = sprintf(
    "%.3f (%.3f, %.3f)",
    exp(beta * iqr_wqs),
    exp((beta - z_crit * se) * iqr_wqs),
    exp((beta + z_crit * se) * iqr_wqs)
  )
)

mixture_effect_copd
```


#3.2 Asthma (inverse winter temperature)

inverse of winter temp
```{r}
u2_df_asthma$inv_u1_gt_t_win_mean_hmg_g <- -u2_df_asthma$u1_gt_t_win_mean_hmg_g
```

```{r}
wqs_vars <- c(
  pm_vars,
  no2_var,
  "inv_u1_gt_t_win_mean_hmg_g"
)
```

```{r}
df           = u2_df_asthma
outcome      = "u2_dt_an_lung_2_fu"

df$outcome_bin <- as.integer(df[[outcome]] == 1)

# Keep only needed columns
wqs_df <- df[, c("outcome_bin", wqs_vars, covars_gwas)]
```

run model
```{r}
fit_wqs_asthma <- gwqs(
  formula = outcome_bin ~ wqs + 
    u1_basist_sex +
    u1_basist_age +
    u1_at_ses_mig_status +
    u1_at_smok_stat_noint +
    u1_at_ses_isced11_3level +
    u1_gt_gisd_rki_m +
    u1_gt_pop_d_ink_a_18 +
    u1_basist_uort_combined,
  mix_name = wqs_vars,
  data = wqs_df,
  q = 10,
  validation = 0.6,
  b = 100,
  family = "binomial",
  seed = 809,
  tables = TRUE
)
```

weights
```{r}
fit_wqs_asthma$final_weights
```

convert results to IQR scale
```{r}
# coefficient table
coef_tab <- summary(fit_wqs_asthma$fit)$coefficients

# extract WQS term
beta <- coef_tab["wqs", "Estimate"]
se   <- coef_tab["wqs", "Std. Error"]

# IQR of observed WQS index
iqr_wqs <- IQR(fit_wqs_asthma$fit$model$wqs, na.rm = TRUE)  #can switch this to 5
#iqr_wqs <- 5

# inference
z_val   <- beta / se
p_value <- 2 * (1 - pnorm(abs(z_val)))
z_crit  <- qnorm(0.975)

# OR per IQR increase in the WQS index
mixture_effect_asthma <- tibble::tibble(
  exposure = "mixture",
  IQR_WQS  = iqr_wqs,
  OR       = exp(beta * iqr_wqs),
  CI_low   = exp((beta - z_crit * se) * iqr_wqs),
  CI_high  = exp((beta + z_crit * se) * iqr_wqs),
  p_value  = p_value,
  OR_CI    = sprintf(
    "%.3f (%.3f, %.3f)",
    exp(beta * iqr_wqs),
    exp((beta - z_crit * se) * iqr_wqs),
    exp((beta + z_crit * se) * iqr_wqs)
  )
)

mixture_effect_asthma
```


#3.3 create WQS plots
```{r}
titles <- c(
  "inv_u1_gt_t_win_mean_hmg_g" = "Inverse Winter\nTemperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2"
)
```

```{r}
p1 <- ggplot(mixture_effect_asthma, aes(x = exposure, y = OR)) +
  geom_point(size = 2, color = "black") +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    width = 0.18,
    color = "black",
    linewidth = 1 
  ) +
  scale_y_continuous(limits = c(0.94, 1.6)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    text = element_text(size = 15),       
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )

p2 <- ggplot(mixture_effect_copd, aes(x = exposure, y = OR)) +
  geom_point(size = 2, color = "black") +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    width = 0.18,
    color = "black",
    linewidth = 1 
  ) +
  scale_y_continuous(limits = c(0.94, 1.6)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    text = element_text(size = 15),         
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )


# convert weights to dataframe
weights_df <- as.data.frame(fit_wqs_asthma$final_weights)
colnames(weights_df) <- c("component", "weight")
weights_df <- weights_df %>%
  arrange(desc(weight))
p3 <- ggplot(weights_df, aes(x = reorder(component, weight), y = weight)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(labels = titles) + 
  labs(
    x = "",
    y = "Weight"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 15),       
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )

weights_df <- as.data.frame(fit_wqs_copd$final_weights)
colnames(weights_df) <- c("component", "weight")
weights_df <- weights_df %>%
  arrange(desc(weight))
p4 <- ggplot(weights_df, aes(x = reorder(component, weight), y = weight)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(labels = titles) + 
  labs(
    x = "",
    y = "Weight"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 15),      
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )


top <- plot_grid(p1, p3, p3, ncol = 3, rel_widths = c(0.6, 0.9, 0.5))
bottom <- plot_grid(p2, p4, p4, ncol = 3, rel_widths = c(0.6, 0.9, 0.5))

final_plot <- plot_grid(top, bottom, ncol = 1)
final_plot

#save as PNG
ggsave(
  "",
  plot = final_plot,
  width = 10,
  height = 6,
  dpi = 1200
)
```



#4 Effect modification analyses
- For Figure 3 and Table S4

list of variables to assess
```{r}
# stratifiers where NA rows should be dropped (not included in the main model so contain missing)
drop_na_stratifiers <- c(
  "u1_at_anthro_bmi_category",
  "u1_d_bb1",
  "u1_at_gpaq_ptotal_quartile",
  "u1_gt_dgu_eu_m",
  "u1_gt_rh_mean_hmg_g")


individual_covars <- list(
  humidity = list(
    title  = "humidity",
    var     = "rh_quart3",
    mainvar = "rh_quart3",
    labels  = c(
      "0–25th"   = "Low (0–25th percentile)",
      "25–75th"  = "Medium (25–75th percentile)",
      "75–100th" = "High (75–100th percentile)"
    )
  ),
  sex = list(
    title  = "Sex",
    var    = "u1_basist_sex",
    mainvar = "u1_basist_sex",
    labels = c(
      "0" = "female",
      "1" = "male"
    )
  ),
  age = list(
    title  = "Age Group",
    var    = "u1_basist_age_category",
    mainvar = "u1_basist_age",
    labels = c(
      "0" = "< 65",
      "1" = ">= 65"
    )
  ),
  obesity = list(
    title  = "Obesity",
    var    = "u1_at_anthro_bmi_category",
    mainvar = "u1_at_anthro_bmi_category",
    labels = c(
      "0" = "no",
      "1" = "yes"
    )
  ),
  smoking = list(
    title = "Smoking Status",
    var = "u1_at_smok_stat_noint",
    mainvar = "u1_at_smok_stat_noint",
    labels = c(
      "0" = "never smoker",
      "1" = "previous smoker",
      "2" = "current smoker"
    )
  ),
  occupational = list(
    title = "Occupational Exposure",
    var = "u1_d_bb1",
    mainvar = "u1_d_bb1",
    labels = c(
      "0" = "no",
      "1" = "yes, wore mask",
      "2" = "yes, did not wear mask"
    )
  ),
  physical_activity = list(
    title  = "Physical Activity",
    var    = "u1_at_gpaq_ptotal_quartile",
    mainvar = "u1_at_gpaq_ptotal_quartile",
    labels = c(
      "0" = "1st quartile",
      "1" = "2nd quartile",
      "2" = "3rd quartile",
      "3" = "4th quartile"
    )
  ),
  urabnization = list(
    title = "Urbanization",
    var = "u1_gt_dgu_eu_m",
    mainvar = "",
    labels = c(
      "0"     = "Rural",
      "1"  = "Suburban",
      "2"     = "Urban"
    )
  )
)
```


IQR scale variables
```{r}
prescale_vars_iqr <- function(df, vars) {
  for (x in vars) {
    iqr_x <- stats::IQR(df[[x]], na.rm = TRUE)
    df[[paste0(x, "_iqr")]] <- df[[x]] / iqr_x
  }
  df
}

for (i in seq_along(scenarios)) {
  scenarios[[i]]$df <- prescale_vars_iqr(scenarios[[i]]$df, all_vars)
}
```

function to run interaction analyses
```{r}
z_975 <- qnorm(0.975)

fit_logistic_interaction <- function(
    df,
    y,
    x,
    covars,
    interaction_var = NULL,
    study_center_random_int = FALSE
) {
  
  scaled_name <- paste0(x, "_iqr")
  
  # enforce factor
  if (!is.null(interaction_var)) {
    df[[interaction_var]] <- factor(df[[interaction_var]])
  }
  
  covars_clean <- setdiff(covars, interaction_var)
  
  fixed_terms <- c(scaled_name, covars_clean)
  
  if (!is.null(interaction_var)) {
    fixed_terms <- c(
      fixed_terms,
      interaction_var,
      paste0(scaled_name, ":", interaction_var)
    )
  }
  
  # MODEL
  if (study_center_random_int) {
    
    form <- reformulate(
      c(fixed_terms, paste0("(1 | ", study_center_var, ")")),
      response = y
    )
    
    fit <- lme4::glmer(
      form,
      data = df,
      family = binomial(),
      control = lme4::glmerControl(
        optimizer = "bobyqa",
        calc.derivs = FALSE,
        optCtrl = list(maxfun = 2e5)
      )
    )
    
  } else {
    
    form <- reformulate(fixed_terms, response = y)
    fit <- glm(form, data = df, family = binomial())
  }
  
  coefs <- summary(fit)$coefficients
  
  # main effect (reference)
  beta_main <- coefs[scaled_name, "Estimate"]
  se_main   <- coefs[scaled_name, "Std. Error"]
  p_main    <- coefs[scaled_name, "Pr(>|z|)"]
  
  OR_no <- exp(beta_main)
  CI_no_low  <- exp(beta_main - z_975 * se_main)
  CI_no_high <- exp(beta_main + z_975 * se_main)
  
  if (is.null(interaction_var)) {
    return(tibble::tibble(
      exposure = x,
      group = "reference",
      OR = OR_no,
      CI_low = CI_no_low,
      CI_high = CI_no_high,
      p_value = p_main,
      p_int_coef = NA_real_
    ))
  }
  
  # multi-level interaction
  vc <- vcov(fit)
  
  int_terms <- grep(
    paste0("^", scaled_name, ":", interaction_var),
    rownames(coefs),
    value = TRUE
  )
  
  # Reference group
  results <- list(
    tibble::tibble(
      exposure = x,
      group = levels(df[[interaction_var]])[1], 
      OR = OR_no,
      CI_low = CI_no_low,
      CI_high = CI_no_high,
      p_value = p_main,
      p_int_coef = NA_real_
    )
  )
  
  for (term_name in int_terms) {
    
    beta_int <- coefs[term_name, "Estimate"]
    se_int   <- coefs[term_name, "Std. Error"]
    p_int    <- coefs[term_name, "Pr(>|z|)"]
    
    covar <- vc[scaled_name, term_name]
    
    beta_total <- beta_main + beta_int
    se_total   <- sqrt(se_main^2 + se_int^2 + 2 * covar)
    
    OR <- exp(beta_total)
    CI_low  <- exp(beta_total - z_975 * se_total)
    CI_high <- exp(beta_total + z_975 * se_total)
    
    level <- sub(
      paste0("^", scaled_name, ":", interaction_var),
      "",
      term_name
    )
    
    results[[length(results) + 1]] <- tibble::tibble(
      exposure = x,
      group = level,
      OR = OR,
      CI_low = CI_low,
      CI_high = CI_high,
      p_value = p_main,
      p_int_coef = p_int
    )
  }
  
  dplyr::bind_rows(results)
}
```

function to run the interaction analyses in a loop
```{r}
run_interaction_models <- function(
    vars,
    df,
    y,
    covars,
    interaction_var,
    mainvar,
    labels
)
{
  
  covars_adj <- setdiff(covars, mainvar)
  
  if (interaction_var %in% drop_na_stratifiers) {
    df <- df[!is.na(df[[interaction_var]]), , drop = FALSE]
  }
  
  # enforce factor
  df[[interaction_var]] <- factor(
    df[[interaction_var]],
    levels = names(labels),
    labels = labels
  )
  
  res <- future.apply::future_lapply(
    vars,
    function(x) {
      fit_logistic_interaction(
        df     = df,
        y      = y,
        x      = x,
        covars = covars_adj,
        interaction_var = interaction_var,
        study_center_random_int = TRUE
      )
    }
  )
  
  dplyr::bind_rows(res) |>
    dplyr::mutate(interaction = interaction_var)
}
```

run the loop
```{r}
all_interaction_results <- lapply(
  individual_covars,
  function(info) {
    lapply(scenarios, function(s) {
      run_interaction_models(
        vars = all_vars,
        df = s$df,
        y = s$outcome,
        covars = covars,
        interaction_var = info$var,
        mainvar = info$mainvar,
        labels = info$labels
      )
    })
  }
)
```

combine results
```{r}
combined_results <- purrr::map_dfr(
  names(individual_covars),
  function(var_name) {
    
    info <- individual_covars[[var_name]]
    
    purrr::map_dfr(seq_along(scenarios), function(i) {
      
      res <- all_interaction_results[[var_name]][[i]]
      if (is.null(res) || nrow(res) == 0) return(NULL)
      
      res %>%
        dplyr::mutate(
          outcome = scenarios[[i]]$outcome,
          interaction = info$title,
          stratum = factor(
            group,
            levels = unname(info$labels)  
          ),
          scenario_title = dplyr::recode(
            scenarios[[i]]$outcome,
            "u2_dt_an_lung_1_fu" = "COPD Incidence",
            "u2_dt_an_lung_2_fu" = "Asthma Incidence"
          ),
          exposure_label = factor(
            dplyr::recode(exposure, !!!labels),
            levels = labels[names(labels) %in% exposure]
          )
        )
    })
  }
)
```

create table-ready results
```{r}
combined_results <- combined_results %>%
  dplyr::mutate(
    stratum = factor(stratum),
    
    # OR (95% CI)
    `OR (95% CI)` = sprintf(
      "%.3f (%.3f–%.3f)",
      OR,
      CI_low,
      CI_high
    ),
    
    # main p-value
    overall_p_value = round(p_value, 3),
    
    # coefficient-level interaction p-value
    p_interaction = round(p_int_coef, 3)
  )
```

save table
```{r}
readr::write_csv(
  combined_results,
  file = ""
)
```


plotting
```{r}
plots_by_stratifier <- combined_results_test %>%
  split(.$interaction) %>%
  purrr::imap(function(df, strat_name) {
    
    ggplot(
      df,
      aes(
        x = exposure_label,
        y = OR,
        color = stratum,
        group = stratum
      )
    ) +
      
      geom_point(
        position = position_dodge(width = 0.7),
        size = 3
      ) +
      
      geom_errorbar(
        aes(ymin = CI_low, ymax = CI_high),
        width = 0.25,
        position = position_dodge(width = 0.7)
      ) +
      
      geom_hline(
        yintercept = 1,
        linetype = "dashed",
        color = "gray40"
      ) +
      
      scale_y_log10(
        breaks = seq(0.8, 1.4, by = 0.1),
        labels = scales::number_format(accuracy = 0.1)
      ) +
      
      facet_wrap(~ scenario_title) +
      
      scale_color_jama(name = strat_name) +
      
      labs(
        title = paste("Stratified by", strat_name),
        x = "",
        y = "Odds Ratio"
      ) +
      
      theme_minimal(base_size = 19) +
      
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title  = element_text(face = "bold"),
        strip.text  = element_text(size = 19, face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.line.y = element_line(color = "gray50", linewidth = 0.5),
        axis.line.x = element_line(color = "gray50", linewidth = 0.5),
        legend.position = "right" 
      )
  })
```

```{r}
purrr::walk(plots_by_stratifier, print)
```



