
This file contains the code to run single-exposure models (GAMMs and logistic regression)

```{r}
library(dplyr)
library(mgcv)
library(purrr)
library(tibble)
library(patchwork)
library(ggplot2)
library(future)
library(future.apply)
library(gWQS)
library(ggsci)
library(cowplot)
library(scales)
```


# Load data
```{r}
#main datasets
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline
```


# Setting up variables

setting up to run code in parallel
```{r}
plan(
  multisession,
  workers = max(1, parallel::detectCores() - 1)
)

options(future.globals.maxSize = 8 * 1024^3)
```

setting up scenarios--one for COPD, one for asthma
```{r}
scenarios <- list(
  copd_incidence = list(df = u2_df_copd, outcome = "u2_dt_an_lung_1_fu", title = "COPD Incidence"),
  asthma_incidence = list(df = u2_df_asthma, outcome = "u2_dt_an_lung_2_fu", title = "Asthma Incidence")
)
```

setting up variables
```{r}
temp_var <- "u1_gt_t_mean_hmg_g"
winter_var <- "u1_gt_t_win_mean_hmg_g"
temp_vars <- c("u1_gt_t_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g")
seasons_order <- c("u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g")
humidity_var <- "u1_gt_rh_mean_hmg_g"

pm_vars <- c("u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g")
no2_var <-"u1_gt_no2_exp_g"
pollutant_vars <- c(pm_vars, no2_var)

all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)

study_center_var <-"u1_basist_uort_combined"
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)

titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)
```




#1. GAMMS with non-linear terms for environmental data
- For Figure 1 and Figure S5

variable set up -- temp & seasons
```{r}
all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)
```

variable set up -- all temperature seasons
```{r}
all_vars <- c(temp_var, seasons_order)
```

function -- final version
```{r}
fit_gamm <- function(
    y,  data,
    x,
    label,
    title,
    covars,
    study_center_random_int = FALSE
) {
  
  #ensure study center is factor (for random intercept)
  if (study_center_random_int) {
    data[[study_center_var]] <- factor(data[[study_center_var]])
  }
  
  #build formula
  covar_string <- paste(covars, collapse = " + ")
  
  re_term <- if (study_center_random_int) {
    paste0(" + s(", study_center_var, ", bs = 're')")
  } else {
    ""
  }
  
  form <- as.formula(
    paste0(
      y,
      " ~ s(", x, ", k=4) + ",
      covar_string,
      re_term
    )
  )
  
  #fit GAM
  fit <- mgcv::gam(
    formula = form,
    data    = data,
    family  = binomial(link = "logit"),
    method  = "REML"
  )
  
  #prediction grid 
  xseq <- seq(
    min(data[[x]], na.rm = TRUE),
    max(data[[x]], na.rm = TRUE),
    length.out = 200
  )
  
  newd <- data.frame(exposure_value = xseq)
  names(newd)[1] <- x
  
  #set covariates to median / mode
  for (v in covars) {
    if (is.numeric(data[[v]])) {
      newd[[v]] <- median(data[[v]], na.rm = TRUE)
    } else {
      mode_val <- names(sort(table(data[[v]]), decreasing = TRUE))[1]
      newd[[v]] <- factor(mode_val, levels = levels(data[[v]]))
    }
  }
  
  if (study_center_random_int) {
    newd[[study_center_var]] <- levels(data[[study_center_var]])[1]
  }
  
  #predict on log-odds scale
  pred <- predict(
    fit,
    newdata = newd,
    se.fit  = TRUE,
    type    = "link",
    exclude = if (study_center_random_int)
      paste0("s(", study_center_var, ")")
    else
      NULL
  )
  
  #reference = 1st percentile
  ref_val <- stats::quantile(data[[x]], 0.01, na.rm = TRUE)
  ref_idx <- which.min(abs(xseq - ref_val))
  
  #return odds ratios
  return(list(
    preds = tibble::tibble(
      exposure_value = xseq,
      OR      = exp(pred$fit - pred$fit[ref_idx]),
      OR_low  = exp((pred$fit - 1.96 * pred$se.fit) - pred$fit[ref_idx]),
      OR_high = exp((pred$fit + 1.96 * pred$se.fit) - pred$fit[ref_idx]),
      exposure = x,
      label    = label,
      title = title
    ),
    fit = fit
  ))
}
```

function to loop through all environmental variables
- 1st-99th percentiles
- standardizing y-axis
- with histogram
```{r}
gamm_singleexposure_outputs <- lapply(
  scenarios,
  function(s) {
    
    plotdata_list <- future.apply::future_lapply(
      all_vars,
      function(x) {
        fit_gamm(
          data  = s$df,
          y     = s$outcome,
          x     = x,
          label = if (x %in% names(labels)) labels[[x]] else x,
          title = if (x %in% names(titles)) titles[[x]] else x,
          covars = covars,
          study_center_random_int = TRUE
        )
      },
      future.seed = TRUE
    )
    
    names(plotdata_list) <- sapply(
      all_vars,
      function(x) {
        if (x %in% names(labels)) labels[[x]] else x
      }
    )
    
    plots <- lapply(
      seq_along(plotdata_list),
      function(i) {
        
        df <- plotdata_list[[i]]$preds
        
        lab   <- df$label[1]
        title <- df$title[1]
        xvar  <- df$exposure[1]
        
        # percentiles from original data
        p1  <- quantile(s$df[[xvar]], 0.01, na.rm = TRUE)
        p99 <- quantile(s$df[[xvar]], 0.99, na.rm = TRUE)
        p5  <- quantile(s$df[[xvar]], 0.05, na.rm = TRUE)
        p95 <- quantile(s$df[[xvar]], 0.95, na.rm = TRUE)
        
        # restrict GAMM curve
        df <- df %>%
          dplyr::filter(
            exposure_value >= p1,
            exposure_value <= p99
          )
        
        # histogram data
        hist_df <- s$df %>%
          dplyr::filter(
            .data[[xvar]] >= p1,
            .data[[xvar]] <= p99
          )
        
        # GAMM plot (top panel)
        p_gamm <- ggplot(
          df,
          aes(x = exposure_value, y = OR)
        ) +
          geom_ribbon(
            aes(
              ymin = OR_low,
              ymax = OR_high
            ),
            fill = "grey50",
            alpha = 0.3
          ) +
          geom_line(
            linewidth = 1.1,
            color = "black"
          ) +
          geom_hline(
            yintercept = 1,
            #linetype = "dashed",
            color = "gray30",
            linewidth = 0.5
          ) +
          geom_vline(
            xintercept = c(p5, p95),
            linetype = "dashed",
            color = "gray50",
            linewidth = 0.5
          ) +
          coord_cartesian(
            xlim = c(p1, p99)
          ) +
          scale_y_log10(
            limits = c(0.45, 2.8)
          ) +
          labs(
            title = title,
            x = NULL,
            y = "Odds Ratio"
          ) +
          #theme_minimal(base_size = 9) +   #for all seasons plot
          theme_minimal(base_size = 12) +
          theme(
            #plot.title = element_text(hjust = 0.5),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.title.x = element_blank(),
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            plot.margin = margin(5, 5, 0, 5),
            axis.line.y = element_line(color = "gray50", linewidth = 0.5),
          )
        
        # Histogram (bottom panel)
        p_hist <- ggplot(
          hist_df,
          aes(x = .data[[xvar]])
        ) +
          geom_histogram(
            bins = 30,
            fill = "grey75",
            colour = "white",
            linewidth = 0.2
          ) +
          geom_vline(
            xintercept = c(p5, p95),
            linetype = "dashed",
            color = "gray50",
            linewidth = 0.5
          ) +
          coord_cartesian(
            xlim = c(p1, p99)
          ) +
          labs(
            x = lab,
            y = "Count"
          ) +
          #theme_minimal(base_size = 6.5) +   #for all seasons plot
          theme_minimal(base_size = 10) +
          theme(
            plot.margin = margin(0, 5, 5, 5),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.line.y = element_line(color = "gray50", linewidth = 0.5),
            axis.line.x = element_line(color = "gray50", linewidth = 0.5),
            #axis.title.y = element_text(size = 7),
            #axis.text.y = element_text(size = 6)
            axis.text.y  = element_blank(),
            axis.ticks.y = element_blank(),
            axis.title.y = element_blank()
          )
        
        # Combine panels
        p_gamm /
          p_hist +
          patchwork::plot_layout(
            heights = c(9, 1)
          )
      }
    )
    
    list(
      plotdata_list = plotdata_list,
      plots = plots
    )
  }
)
```


plot
```{r}
scenario_boards <- lapply(names(gamm_singleexposure_outputs), function(scn_name) {
  scn   <- gamm_singleexposure_outputs[[scn_name]]
  plots <- scn$plots
  
  board <- wrap_plots(
    plots,
    ncol = 5,
    nrow = 3
  )
  
  board
})

names(scenario_boards) <- names(gamm_singleexposure_outputs)
```

```{r}
scenario_boards$asthma_incidence
scenario_boards$copd_incidence
```


save Figure 1
```{r}
ggsave(
  "",
  plot = scenario_boards$asthma_incidence,
  width = 8,
  height = 6,
  dpi = 1200
)

ggsave(
  "",
  plot = scenario_boards$copd_incidence,
  width = 8,
  height = 6,
  dpi = 1200
)
```

save all seasons plot (Figure S5)
```{r}
ggsave(
  "",
  plot = scenario_boards$asthma_incidence,
  width = 13,
  height = 6,
  dpi = 1200
)

ggsave(
  "",
  plot = scenario_boards$copd_incidence,
  width = 13,
  height = 6,
  dpi = 1200
)
```



#2 adjusted logistic regression with linear term to get ORs
- For Table 2

main function - scale factor is IQR
```{r}
fit_logistic_adjusted <- function(
    df,
    y,
    x,
    covars,
    study_center_random_int = FALSE
) {
  
  # choose scale factor
  # SD
  #sd_x <- stats::sd(df[[x]], na.rm = TRUE)
  #scaled_name <- paste0(x, "_sd")
  #df[[scaled_name]] <- df[[x]] / sd_x
  
  #IQR
  iqr_x <- stats::IQR(df[[x]], na.rm = TRUE)
  scaled_name <- paste0(x, "_iqr")
  df[[scaled_name]] <- df[[x]] / iqr_x
  
  fixed_terms <- c(scaled_name, covars)
  
  
  if (study_center_random_int) {
    fixed_part <- paste(fixed_terms, collapse = " + ")
    rand_part  <- paste0("(1 | ", study_center_var, ")")
    
    form <- as.formula(
      paste(y, "~", fixed_part, "+", rand_part)
    )
    fit <- lme4::glmer(
      form,
      data    = df,
      family  = binomial(),
      control = lme4::glmerControl(
        optimizer   = "bobyqa",
        calc.derivs = FALSE
      )
    )
    
    summ <- summary(fit)$coefficients
    
  } else {
    form <- reformulate(fixed_terms, response = y)
    
    fit <- glm(
      form,
      data   = df,
      family = binomial()
    )
    
    summ <- summary(fit)$coefficients
  }
  
  beta <- summ[scaled_name, "Estimate"]
  se   <- summ[scaled_name, "Std. Error"]
  z    <- qnorm(0.975)
  
  CI_low  <- beta - z * se
  CI_high <- beta + z * se
  p_value <- summ[scaled_name, "Pr(>|z|)"]
  
  OR_val <- exp(beta)
  CI_l   <- exp(CI_low)
  CI_h   <- exp(CI_high)
  
  tibble(
    exposure = x,
    OR       = OR_val,
    CI_low   = CI_l,
    CI_high  = CI_h,
    OR_CI    = sprintf("%.3f (%.3f, %.3f)", OR_val, CI_l, CI_h),
    p_value  = round(p_value, 3)
  )
}
```

new labels
```{r}
titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual\nTemperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer\nTemperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall\nTemperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring\nTemperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)"
)

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual\nTemperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer\nTemperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall\nTemperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring\nTemperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)"
)
```

loop with only final covariates to create Table 2
```{r}
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")
```

loop
```{r}
adjusted_ors2 <- lapply(
  scenarios,
  function(s) {
    
    # Parallelize the model fits (main bottleneck)
    or_table_scaled <- future.apply::future_lapply(
      all_vars,
      function(x) {
        fit_logistic_adjusted(
          df     = s$df,
          y      = s$outcome,
          x      = x,
          covars = covars,
          study_center_random_int = TRUE
        )
      }
    ) |>
      dplyr::bind_rows()
    
    # clean up tables for plotting
    plot_df <- or_table_scaled |>
      dplyr::mutate(
        pretty = factor(labels[exposure], levels = labels[exposure])
      )
    
    #plot
    plt <- ggplot(plot_df, aes(x = pretty, y = OR)) +
      geom_point(size = 3, color = "black") +
      geom_errorbar(
        aes(ymin = CI_low, ymax = CI_high),
        width = 0.12,
        color = "black"
      ) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
      scale_y_log10() +
      labs(
        title = s$title,
        x = "",
        y = "Odds Ratio"
      ) +
      theme_minimal(base_size = 16) +
      theme(
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        axis.text.y = element_text(size = 13),
        plot.title  = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.line.y = element_line(color = "gray50", linewidth = 0.5),
        axis.line.x = element_line(color = "gray50", linewidth = 0.5)
      )
    
    list(
      or_table = or_table_scaled,
      plot_df  = plot_df,
      plot     = plt
    )
  }
)
```

print
```{r}
adjusted_ors2$asthma_incidence$plot
adjusted_ors2$copd_incidence$plot

asthma_table <- adjusted_ors2$asthma_incidence$or_table
copd_table <- adjusted_ors2$copd_incidence$or_table
```

save
```{r}
write.csv(asthma_table, "", row.names = FALSE)
write.csv(copd_table,   "",   row.names = FALSE)
```
