
This file contains all the code to run the sensitivity analyses included in the paper

```{r}
library(dplyr)
library(tibble)
library(future)
library(future.apply)
library(ggplot2)
library(ggsci)
```


# load data
```{r}
#complete case datasets used in the main models
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline

#expanded covariate complete case datasets
u2_df_asthma_expcovars   <- readRDS("")
u2_df_copd_expcovars   <- readRDS("")

#imputed datasets
u2_df_asthma_imp   <- readRDS("")
u2_df_copd_imp   <- readRDS("")
```



# setting up variables

parallelism
```{r}
plan(
  multisession,
  workers = max(1, parallel::detectCores() - 1)
)

options(future.globals.maxSize = 8 * 1024^3)
```

setting up scenarios
```{r}
scenarios <- list(
  copd_incidence = list(df = u2_df_copd, outcome = "u2_dt_an_lung_1_fu", title = "COPD Incidence"),
  asthma_incidence = list(df = u2_df_asthma, outcome = "u2_dt_an_lung_2_fu", title = "Asthma Incidence")
)
```

setting up variables
```{r}
temp_var <- "u1_gt_t_mean_hmg_g"
winter_var <- "u1_gt_t_win_mean_hmg_g"
temp_vars <- c("u1_gt_t_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g")
seasons_order <- c("u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g")
humidity_var <- "u1_gt_rh_mean_hmg_g"

pm_vars <- c("u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g")
no2_var <-"u1_gt_no2_exp_g"
pollutant_vars <- c(pm_vars, no2_var)

all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)

study_center_var <-"u1_basist_uort_combined"
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")

titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)
```



# Set up the main funciton for adjusted logistic regression with linear term to get ORs

main function - scale factor is IQR
```{r}
fit_logistic_adjusted <- function(
    df,
    y,
    x,
    covars,
    study_center_random_int = FALSE
) {
  
  # choose scale factor
  # SD
  #sd_x <- stats::sd(df[[x]], na.rm = TRUE)
  #scaled_name <- paste0(x, "_sd")
  #df[[scaled_name]] <- df[[x]] / sd_x
  
  #IQR
  iqr_x <- stats::IQR(df[[x]], na.rm = TRUE)
  scaled_name <- paste0(x, "_iqr")
  df[[scaled_name]] <- df[[x]] / iqr_x
  
  fixed_terms <- c(scaled_name, covars)
  
  
  if (study_center_random_int) {
    fixed_part <- paste(fixed_terms, collapse = " + ")
    rand_part  <- paste0("(1 | ", study_center_var, ")")
    
    form <- as.formula(
      paste(y, "~", fixed_part, "+", rand_part)
    )
    fit <- lme4::glmer(
      form,
      data    = df,
      family  = binomial(),
      control = lme4::glmerControl(
        optimizer   = "bobyqa",
        calc.derivs = FALSE
      )
    )
    
    summ <- summary(fit)$coefficients
    
  } else {
    form <- reformulate(fixed_terms, response = y)
    
    fit <- glm(
      form,
      data   = df,
      family = binomial()
    )
    
    summ <- summary(fit)$coefficients
  }
  
  beta <- summ[scaled_name, "Estimate"]
  se   <- summ[scaled_name, "Std. Error"]
  z    <- qnorm(0.975)
  
  CI_low  <- beta - z * se
  CI_high <- beta + z * se
  p_value <- summ[scaled_name, "Pr(>|z|)"]
  
  OR_val <- exp(beta)
  CI_l   <- exp(CI_low)
  CI_h   <- exp(CI_high)
  
  tibble(
    exposure = x,
    OR       = OR_val,
    CI_low   = CI_l,
    CI_high  = CI_h,
    OR_CI    = sprintf("%.3f (%.3f, %.3f)", OR_val, CI_l, CI_h),
    p_value  = round(p_value, 3)
  )
}
```


#1. Main Sensitivity Analysis: testing different covariates, imputation
- Created for Figure S6 and Table S2

list of different covariates
```{r}
all_vars <- c(temp_var, seasons_order, pollutant_vars)

covars_list <- list(
  baseline = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level"
  ),
  GSID = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m"
  ),
  pop_d = c(               
    "u1_basist_sex",
    "u2_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m",
    "u1_gt_pop_d_ink_a_18"
  ),
  #same as population density, as the study center will be added as a random intercept
  main_model = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m",
    "u1_gt_pop_d_ink_a_18"
  ),
  humidity = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m",
    "u1_gt_pop_d_ink_a_18",
    "u1_gt_rh_mean_hmg_g"
  ),
  expanded_covars = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m",
    "u1_gt_pop_d_ink_a_18",
    "u1_at_gpaq_ptotal_quartile",
    "u1_d_bb1",
    "u1_at_anthro_bmi_eig"
  ),
  imputed = c(
    "u1_basist_sex",
    "u1_basist_age",
    "u1_at_ses_mig_status",
    "u1_at_smok_stat_noint",
    "u1_at_ses_isced11_3level",
    "u1_gt_gisd_rki_m",
    "u1_gt_pop_d_ink_a_18",
    "u1_at_gpaq_ptotal_quartile",
    "u1_d_bb1",
    "u1_at_anthro_bmi_eig"
  )
)

model_order <- c(
  "baseline",
  "GSID",
  "pop_d",
  "main_model",
  "humidity",
  "expanded_covars",
  "imputed"
)

model_labels <- c(
  baseline         = "M1",
  GSID             = "M2",
  pop_d            = "M3",
  main_model       = "M4 (Main Model)",
  humidity         = "M5",
  expanded_covars  = "M6",
  imputed          = "M7"
)
```

loop through all models, environmental variables
```{r}
adjusted_ors_multiplecovars <- lapply(
  scenarios,
  function(s) {
    
    # fit models for each covariate set
    all_models <- lapply(
      names(covars_list),
      function(model_name) {
        
        covars <- covars_list[[model_name]]
        
        future.apply::future_lapply(
          all_vars,
          function(x) {
            
            # select the right df
            df_use <- if (model_name == "expanded_covars") {
              
              if (s$title == "COPD Incidence") {
                u2_df_copd_expcovars
              } else if (s$title == "Asthma Incidence") {
                u2_df_asthma_expcovars
              } else {
                s$df
              }
              
            } else if (model_name == "imputed") {
              
              if (s$title == "COPD Incidence") {
                u2_df_copd_imp
              } else if (s$title == "Asthma Incidence") {
                u2_df_asthma_imp
              } else {
                s$df
              }
              
            } else {
              s$df
            }
            
            fit_logistic_adjusted(
              df     = df_use,
              y      = s$outcome,
              x      = x,
              covars = covars,
              study_center_random_int = model_name %in% c("main_model","expanded_covars", "imputed", "humidity")
              
            )
          }
        ) |>
          dplyr::bind_rows() |>
          dplyr::mutate(model = model_name)
      }
    ) |>
      dplyr::bind_rows()
    
    # plotting
    plot_df <- all_models |>
      dplyr::mutate(
        exposure_pretty = factor(
          titles[exposure],
          levels = titles[all_vars] 
        ),
        model = factor(
          model,
          levels = names(model_labels),
          labels = model_labels
        )
      )
    
    plt <- ggplot(
      plot_df,
      aes(
        x     = exposure_pretty,
        y     = OR,
        color = model,
        group = model
      )
    ) +
      geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
      geom_point(
        position = position_dodge(width = 0.5),
        size = 3,
        shape = 16
      ) +
      geom_errorbar(
        aes(
          ymin = CI_low,
          ymax = CI_high,
          linewidth = model == "Main Model\n(Study Center)"
        ),
        position = position_dodge(width = 0.5),
        width = 0.42
      ) +
      scale_linewidth_manual(
        values = c(`TRUE` = 1.5, `FALSE` = 0.8),
        guide = "none"
      ) +
      scale_y_log10() +
      scale_color_jama(name = "Model") +
      guides(
        color = guide_legend(
          keyheight = grid::unit(1.1, "cm"),
          keywidth  = grid::unit(0.8, "cm"),
          label.hjust = 0
        )
      ) +
      labs(
        title = s$title,
        x = "",
        y = "Odds Ratio"
      ) +
      theme_minimal(base_size = 20) +
      theme(
        #legend.position = "right",
        legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.line.y = element_line(color = "gray50", linewidth = 0.7),
        axis.line.x = element_line(color = "gray50", linewidth = 0.7),
        axis.ticks = element_line(color = "gray50", linewidth = 0.7),
        axis.ticks.length = unit(3, "pt")
      )
    
    
    list(
      or_table = all_models,
      plot_df  = plot_df,
      plot     = plt
    )
  }
)

# preserve scenario names
names(adjusted_ors_multiplecovars) <- names(scenarios)
```

print
```{r}
adjusted_ors_multiplecovars$asthma_incidence$plot
adjusted_ors_multiplecovars$copd_incidence$plot

asthma_ors <- adjusted_ors_multiplecovars$asthma_incidence$or_table
copd_ors <- adjusted_ors_multiplecovars$copd_incidence$or_table
```

save figures as PNGs
```{r}
ggsave(
  "",
  plot = adjusted_ors_multiplecovars$asthma_incidence$plot,
  width = 10,
  height = 6,
  dpi = 1200
)

ggsave(
  "",
  plot = adjusted_ors_multiplecovars$copd_incidence$plot,
  width = 10,
  height = 6,
  dpi = 1200
)
```


save dfs as csvs
```{r}
write.csv(
  adjusted_ors_multiplecovars$asthma_incidence$or_table,
  "asthma_ors.csv",
  row.names = FALSE
)

write.csv(
  adjusted_ors_multiplecovars$copd_incidence$or_table,
  "copd_ors.csv",
  row.names = FALSE
)
```

