
This file contains the code to run joint exposure models (weighted quantile sum)

```{r}
library(dplyr)
library(mgcv)
library(purrr)
library(tibble)
library(patchwork)
library(ggplot2)
library(future)
library(future.apply)
library(gWQS)
library(ggsci)
library(cowplot)
library(scales)
```


# Load data
```{r}
#main datasets
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline
```


# Setting up variables

setting up to run code in parallel
```{r}
plan(
  multisession,
  workers = max(1, parallel::detectCores() - 1)
)

options(future.globals.maxSize = 8 * 1024^3)
```

setting up scenarios--one for COPD, one for asthma
```{r}
scenarios <- list(
  copd_incidence = list(df = u2_df_copd, outcome = "u2_dt_an_lung_1_fu", title = "COPD Incidence"),
  asthma_incidence = list(df = u2_df_asthma, outcome = "u2_dt_an_lung_2_fu", title = "Asthma Incidence")
)
```

setting up variables
```{r}
temp_var <- "u1_gt_t_mean_hmg_g"
winter_var <- "u1_gt_t_win_mean_hmg_g"
temp_vars <- c("u1_gt_t_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g")
seasons_order <- c("u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g")
humidity_var <- "u1_gt_rh_mean_hmg_g"

pm_vars <- c("u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g")
no2_var <-"u1_gt_no2_exp_g"
pollutant_vars <- c(pm_vars, no2_var)

all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)

study_center_var <-"u1_basist_uort_combined"
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)

titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)
```



# grouped weighted quantile sum regression
- For Figure 2 and Table S3

variables
- note that the gwas does not allow a random intercept so we include the study center as a covariate
```{r}
covars_gwas <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18",
  "u1_basist_uort_combined")
```


#1 COPD (winter temperature)
```{r}
wqs_vars <- c(
  "u1_gt_t_win_mean_hmg_g",
  pm_vars,
  no2_var
)
```

```{r}
df           = u2_df_copd
outcome      = "u2_dt_an_lung_1_fu"

df$outcome_bin <- as.integer(df[[outcome]] == 1)

# Keep only needed columns
wqs_df <- df[, c("outcome_bin", wqs_vars, covars_gwas)]
```

run model
```{r}
fit_wqs_copd <- gwqs(
  formula = outcome_bin ~ wqs + 
    u1_basist_sex +
    u1_basist_age +
    u1_at_ses_mig_status +
    u1_at_smok_stat_noint +
    u1_at_ses_isced11_3level +
    u1_gt_gisd_rki_m +
    u1_gt_pop_d_ink_a_18 +
    u1_basist_uort_combined,
  mix_name = wqs_vars,
  data = wqs_df,
  q = 10,
  validation = 0.6,
  b = 100,
  family = "binomial",
  seed = 809,
  tables = TRUE
)
```


get the weights for each environmental variable
```{r}
fit_wqs_copd$final_weights
```

convert results to IQR scale
```{r}
# coefficient table
coef_tab <- summary(fit_wqs_copd$fit)$coefficients

# extract WQS term
beta <- coef_tab["wqs", "Estimate"]
se   <- coef_tab["wqs", "Std. Error"]

# IQR of observed WQS index
iqr_wqs <- IQR(fit_wqs_copd$fit$model$wqs, na.rm = TRUE)

# inference
z_val   <- beta / se
p_value <- 2 * (1 - pnorm(abs(z_val)))
z_crit  <- qnorm(0.975)

# OR per IQR increase in the WQS index
mixture_effect_copd <- tibble::tibble(
  exposure = "mixture",
  IQR_WQS  = iqr_wqs,
  OR       = exp(beta * iqr_wqs),
  CI_low   = exp((beta - z_crit * se) * iqr_wqs),
  CI_high  = exp((beta + z_crit * se) * iqr_wqs),
  p_value  = p_value,
  OR_CI    = sprintf(
    "%.3f (%.3f, %.3f)",
    exp(beta * iqr_wqs),
    exp((beta - z_crit * se) * iqr_wqs),
    exp((beta + z_crit * se) * iqr_wqs)
  )
)

mixture_effect_copd
```


#2 Asthma (inverse winter temperature)

inverse of winter temp
```{r}
u2_df_asthma$inv_u1_gt_t_win_mean_hmg_g <- -u2_df_asthma$u1_gt_t_win_mean_hmg_g
```

```{r}
wqs_vars <- c(
  pm_vars,
  no2_var,
  "inv_u1_gt_t_win_mean_hmg_g"
)
```

```{r}
df           = u2_df_asthma
outcome      = "u2_dt_an_lung_2_fu"

df$outcome_bin <- as.integer(df[[outcome]] == 1)

# Keep only needed columns
wqs_df <- df[, c("outcome_bin", wqs_vars, covars_gwas)]
```

run model
```{r}
fit_wqs_asthma <- gwqs(
  formula = outcome_bin ~ wqs + 
    u1_basist_sex +
    u1_basist_age +
    u1_at_ses_mig_status +
    u1_at_smok_stat_noint +
    u1_at_ses_isced11_3level +
    u1_gt_gisd_rki_m +
    u1_gt_pop_d_ink_a_18 +
    u1_basist_uort_combined,
  mix_name = wqs_vars,
  data = wqs_df,
  q = 10,
  validation = 0.6,
  b = 100,
  family = "binomial",
  seed = 809,
  tables = TRUE
)
```

weights
```{r}
fit_wqs_asthma$final_weights
```

convert results to IQR scale
```{r}
# coefficient table
coef_tab <- summary(fit_wqs_asthma$fit)$coefficients

# extract WQS term
beta <- coef_tab["wqs", "Estimate"]
se   <- coef_tab["wqs", "Std. Error"]

# IQR of observed WQS index
iqr_wqs <- IQR(fit_wqs_asthma$fit$model$wqs, na.rm = TRUE)  #can switch this to 5
#iqr_wqs <- 5

# inference
z_val   <- beta / se
p_value <- 2 * (1 - pnorm(abs(z_val)))
z_crit  <- qnorm(0.975)

# OR per IQR increase in the WQS index
mixture_effect_asthma <- tibble::tibble(
  exposure = "mixture",
  IQR_WQS  = iqr_wqs,
  OR       = exp(beta * iqr_wqs),
  CI_low   = exp((beta - z_crit * se) * iqr_wqs),
  CI_high  = exp((beta + z_crit * se) * iqr_wqs),
  p_value  = p_value,
  OR_CI    = sprintf(
    "%.3f (%.3f, %.3f)",
    exp(beta * iqr_wqs),
    exp((beta - z_crit * se) * iqr_wqs),
    exp((beta + z_crit * se) * iqr_wqs)
  )
)

mixture_effect_asthma
```


#3 create WQS plots
```{r}
titles <- c(
  "inv_u1_gt_t_win_mean_hmg_g" = "Inverse Winter\nTemperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter\nTemperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2"
)
```

```{r}
p1 <- ggplot(mixture_effect_asthma, aes(x = exposure, y = OR)) +
  geom_point(size = 2, color = "black") +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    width = 0.18,
    color = "black",
    linewidth = 1 
  ) +
  scale_y_continuous(limits = c(0.94, 1.6)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    text = element_text(size = 15),       
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )

p2 <- ggplot(mixture_effect_copd, aes(x = exposure, y = OR)) +
  geom_point(size = 2, color = "black") +
  geom_errorbar(
    aes(ymin = CI_low, ymax = CI_high),
    width = 0.18,
    color = "black",
    linewidth = 1 
  ) +
  scale_y_continuous(limits = c(0.94, 1.6)) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  labs(
    x = "",
    y = "Odds Ratio"
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    text = element_text(size = 15),         
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )


# convert weights to dataframe
weights_df <- as.data.frame(fit_wqs_asthma$final_weights)
colnames(weights_df) <- c("component", "weight")
weights_df <- weights_df %>%
  arrange(desc(weight))
p3 <- ggplot(weights_df, aes(x = reorder(component, weight), y = weight)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(labels = titles) + 
  labs(
    x = "",
    y = "Weight"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 15),       
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )

weights_df <- as.data.frame(fit_wqs_copd$final_weights)
colnames(weights_df) <- c("component", "weight")
weights_df <- weights_df %>%
  arrange(desc(weight))
p4 <- ggplot(weights_df, aes(x = reorder(component, weight), y = weight)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(labels = titles) + 
  labs(
    x = "",
    y = "Weight"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 15),      
    axis.title = element_text(size = 13),
    axis.text = element_text(size = 13),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    axis.line.y = element_line(color = "gray50", linewidth = 0.5),
    axis.line.x = element_line(color = "gray50", linewidth = 0.5)
  )


top <- plot_grid(p1, p3, p3, ncol = 3, rel_widths = c(0.6, 0.9, 0.5))
bottom <- plot_grid(p2, p4, p4, ncol = 3, rel_widths = c(0.6, 0.9, 0.5))

final_plot <- plot_grid(top, bottom, ncol = 1)
final_plot

#save as PNG
ggsave(
  "",
  plot = final_plot,
  width = 10,
  height = 6,
  dpi = 1200
)
```