
This file contains the code to run effect modification analyses for individual and neighborhood-level vulnerabilities.

```{r}
library(dplyr)
library(mgcv)
library(purrr)
library(tibble)
library(patchwork)
library(ggplot2)
library(future)
library(future.apply)
library(gWQS)
library(ggsci)
library(cowplot)
library(scales)
```


# Load data
```{r}
#main datasets
u2_df_asthma <- readRDS("") #5-year follow-up df, excluding those with asthma at baseline
u2_df_copd   <- readRDS("") #5-year follow-up df, excluding those with COPD at baseline
```


# Setting up variables

setting up to run code in parallel
```{r}
plan(
  multisession,
  workers = max(1, parallel::detectCores() - 1)
)

options(future.globals.maxSize = 8 * 1024^3)
```

setting up scenarios--one for COPD, one for asthma
```{r}
scenarios <- list(
  copd_incidence = list(df = u2_df_copd, outcome = "u2_dt_an_lung_1_fu", title = "COPD Incidence"),
  asthma_incidence = list(df = u2_df_asthma, outcome = "u2_dt_an_lung_2_fu", title = "Asthma Incidence")
)
```

setting up variables
```{r}
temp_var <- "u1_gt_t_mean_hmg_g"
winter_var <- "u1_gt_t_win_mean_hmg_g"
temp_vars <- c("u1_gt_t_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g")
seasons_order <- c("u1_gt_t_fal_mean_hmg_g", "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", "u1_gt_t_smr_mean_hmg_g")
humidity_var <- "u1_gt_rh_mean_hmg_g"

pm_vars <- c("u1_gt_pm25_exp_g", "u1_gt_pm10_exp_g")
no2_var <-"u1_gt_no2_exp_g"
pollutant_vars <- c(pm_vars, no2_var)

all_vars <- c(temp_var, "u1_gt_t_win_mean_hmg_g", "u1_gt_t_spr_mean_hmg_g", pollutant_vars)

study_center_var <-"u1_basist_uort_combined"
covars <- c(
  "u1_basist_sex",
  "u1_basist_age",
  "u1_at_ses_mig_status",
  "u1_at_smok_stat_noint",
  "u1_at_ses_isced11_3level",
  "u1_gt_gisd_rki_m",
  "u1_gt_pop_d_ink_a_18")

labels <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature (°C)",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature (°C)",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature (°C)",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature (°C)",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature (°C)",
  "u1_gt_pm25_exp_g"   = "PM2.5 (µg/m³)",
  "u1_gt_pm10_exp_g"   = "PM10 (µg/m³)",
  "u1_gt_no2_exp_g"    = "NO2 (µg/m³)",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)

titles <- c(
  "u1_gt_t_mean_hmg_g" = "Annual Temperature",
  "u1_gt_t_win_mean_hmg_g" = "Winter Temperature",
  "u1_gt_t_smr_mean_hmg_g" = "Summer Temperature",
  "u1_gt_t_fal_mean_hmg_g" = "Fall Temperature",
  "u1_gt_t_spr_mean_hmg_g" = "Spring Temperature",
  "u1_gt_pm25_exp_g"   = "PM2.5",
  "u1_gt_pm10_exp_g"   = "PM10",
  "u1_gt_no2_exp_g"    = "NO2",
  "u1_gt_gisd_rki_m" = "GISD",
  "u1_gt_rh_mean_hmg_g" = "Relative Humidity"
)
```


# Effect modification analyses
- For Figure 3 and Table S4

list of variables to assess
```{r}
# stratifiers where NA rows should be dropped (not included in the main model so contain missing)
drop_na_stratifiers <- c(
  "u1_at_anthro_bmi_category",
  "u1_d_bb1",
  "u1_at_gpaq_ptotal_quartile",
  "u1_gt_dgu_eu_m",
  "u1_gt_rh_mean_hmg_g")


individual_covars <- list(
  humidity = list(
    title  = "humidity",
    var     = "rh_quart3",
    mainvar = "rh_quart3",
    labels  = c(
      "0–25th"   = "Low (0–25th percentile)",
      "25–75th"  = "Medium (25–75th percentile)",
      "75–100th" = "High (75–100th percentile)"
    )
  ),
  sex = list(
    title  = "Sex",
    var    = "u1_basist_sex",
    mainvar = "u1_basist_sex",
    labels = c(
      "0" = "female",
      "1" = "male"
    )
  ),
  age = list(
    title  = "Age Group",
    var    = "u1_basist_age_category",
    mainvar = "u1_basist_age",
    labels = c(
      "0" = "< 65",
      "1" = ">= 65"
    )
  ),
  obesity = list(
    title  = "Obesity",
    var    = "u1_at_anthro_bmi_category",
    mainvar = "u1_at_anthro_bmi_category",
    labels = c(
      "0" = "no",
      "1" = "yes"
    )
  ),
  smoking = list(
    title = "Smoking Status",
    var = "u1_at_smok_stat_noint",
    mainvar = "u1_at_smok_stat_noint",
    labels = c(
      "0" = "never smoker",
      "1" = "previous smoker",
      "2" = "current smoker"
    )
  ),
  occupational = list(
    title = "Occupational Exposure",
    var = "u1_d_bb1",
    mainvar = "u1_d_bb1",
    labels = c(
      "0" = "no",
      "1" = "yes, wore mask",
      "2" = "yes, did not wear mask"
    )
  ),
  physical_activity = list(
    title  = "Physical Activity",
    var    = "u1_at_gpaq_ptotal_quartile",
    mainvar = "u1_at_gpaq_ptotal_quartile",
    labels = c(
      "0" = "1st quartile",
      "1" = "2nd quartile",
      "2" = "3rd quartile",
      "3" = "4th quartile"
    )
  ),
  urabnization = list(
    title = "Urbanization",
    var = "u1_gt_dgu_eu_m",
    mainvar = "",
    labels = c(
      "0"     = "Rural",
      "1"  = "Suburban",
      "2"     = "Urban"
    )
  )
)
```


IQR scale variables
```{r}
prescale_vars_iqr <- function(df, vars) {
  for (x in vars) {
    iqr_x <- stats::IQR(df[[x]], na.rm = TRUE)
    df[[paste0(x, "_iqr")]] <- df[[x]] / iqr_x
  }
  df
}

for (i in seq_along(scenarios)) {
  scenarios[[i]]$df <- prescale_vars_iqr(scenarios[[i]]$df, all_vars)
}
```

function to run interaction analyses
```{r}
z_975 <- qnorm(0.975)

fit_logistic_interaction <- function(
    df,
    y,
    x,
    covars,
    interaction_var = NULL,
    study_center_random_int = FALSE
) {
  
  scaled_name <- paste0(x, "_iqr")
  
  # enforce factor
  if (!is.null(interaction_var)) {
    df[[interaction_var]] <- factor(df[[interaction_var]])
  }
  
  covars_clean <- setdiff(covars, interaction_var)
  
  fixed_terms <- c(scaled_name, covars_clean)
  
  if (!is.null(interaction_var)) {
    fixed_terms <- c(
      fixed_terms,
      interaction_var,
      paste0(scaled_name, ":", interaction_var)
    )
  }
  
  # MODEL
  if (study_center_random_int) {
    
    form <- reformulate(
      c(fixed_terms, paste0("(1 | ", study_center_var, ")")),
      response = y
    )
    
    fit <- lme4::glmer(
      form,
      data = df,
      family = binomial(),
      control = lme4::glmerControl(
        optimizer = "bobyqa",
        calc.derivs = FALSE,
        optCtrl = list(maxfun = 2e5)
      )
    )
    
  } else {
    
    form <- reformulate(fixed_terms, response = y)
    fit <- glm(form, data = df, family = binomial())
  }
  
  coefs <- summary(fit)$coefficients
  
  # main effect (reference)
  beta_main <- coefs[scaled_name, "Estimate"]
  se_main   <- coefs[scaled_name, "Std. Error"]
  p_main    <- coefs[scaled_name, "Pr(>|z|)"]
  
  OR_no <- exp(beta_main)
  CI_no_low  <- exp(beta_main - z_975 * se_main)
  CI_no_high <- exp(beta_main + z_975 * se_main)
  
  if (is.null(interaction_var)) {
    return(tibble::tibble(
      exposure = x,
      group = "reference",
      OR = OR_no,
      CI_low = CI_no_low,
      CI_high = CI_no_high,
      p_value = p_main,
      p_int_coef = NA_real_
    ))
  }
  
  # multi-level interaction
  vc <- vcov(fit)
  
  int_terms <- grep(
    paste0("^", scaled_name, ":", interaction_var),
    rownames(coefs),
    value = TRUE
  )
  
  # Reference group
  results <- list(
    tibble::tibble(
      exposure = x,
      group = levels(df[[interaction_var]])[1], 
      OR = OR_no,
      CI_low = CI_no_low,
      CI_high = CI_no_high,
      p_value = p_main,
      p_int_coef = NA_real_
    )
  )
  
  for (term_name in int_terms) {
    
    beta_int <- coefs[term_name, "Estimate"]
    se_int   <- coefs[term_name, "Std. Error"]
    p_int    <- coefs[term_name, "Pr(>|z|)"]
    
    covar <- vc[scaled_name, term_name]
    
    beta_total <- beta_main + beta_int
    se_total   <- sqrt(se_main^2 + se_int^2 + 2 * covar)
    
    OR <- exp(beta_total)
    CI_low  <- exp(beta_total - z_975 * se_total)
    CI_high <- exp(beta_total + z_975 * se_total)
    
    level <- sub(
      paste0("^", scaled_name, ":", interaction_var),
      "",
      term_name
    )
    
    results[[length(results) + 1]] <- tibble::tibble(
      exposure = x,
      group = level,
      OR = OR,
      CI_low = CI_low,
      CI_high = CI_high,
      p_value = p_main,
      p_int_coef = p_int
    )
  }
  
  dplyr::bind_rows(results)
}
```

function to run the interaction analyses in a loop
```{r}
run_interaction_models <- function(
    vars,
    df,
    y,
    covars,
    interaction_var,
    mainvar,
    labels
)
{
  
  covars_adj <- setdiff(covars, mainvar)
  
  if (interaction_var %in% drop_na_stratifiers) {
    df <- df[!is.na(df[[interaction_var]]), , drop = FALSE]
  }
  
  # enforce factor
  df[[interaction_var]] <- factor(
    df[[interaction_var]],
    levels = names(labels),
    labels = labels
  )
  
  res <- future.apply::future_lapply(
    vars,
    function(x) {
      fit_logistic_interaction(
        df     = df,
        y      = y,
        x      = x,
        covars = covars_adj,
        interaction_var = interaction_var,
        study_center_random_int = TRUE
      )
    }
  )
  
  dplyr::bind_rows(res) |>
    dplyr::mutate(interaction = interaction_var)
}
```

run the loop
```{r}
all_interaction_results <- lapply(
  individual_covars,
  function(info) {
    lapply(scenarios, function(s) {
      run_interaction_models(
        vars = all_vars,
        df = s$df,
        y = s$outcome,
        covars = covars,
        interaction_var = info$var,
        mainvar = info$mainvar,
        labels = info$labels
      )
    })
  }
)
```

combine results
```{r}
combined_results <- purrr::map_dfr(
  names(individual_covars),
  function(var_name) {
    
    info <- individual_covars[[var_name]]
    
    purrr::map_dfr(seq_along(scenarios), function(i) {
      
      res <- all_interaction_results[[var_name]][[i]]
      if (is.null(res) || nrow(res) == 0) return(NULL)
      
      res %>%
        dplyr::mutate(
          outcome = scenarios[[i]]$outcome,
          interaction = info$title,
          stratum = factor(
            group,
            levels = unname(info$labels)  
          ),
          scenario_title = dplyr::recode(
            scenarios[[i]]$outcome,
            "u2_dt_an_lung_1_fu" = "COPD Incidence",
            "u2_dt_an_lung_2_fu" = "Asthma Incidence"
          ),
          exposure_label = factor(
            dplyr::recode(exposure, !!!labels),
            levels = labels[names(labels) %in% exposure]
          )
        )
    })
  }
)
```

create table-ready results
```{r}
combined_results <- combined_results %>%
  dplyr::mutate(
    stratum = factor(stratum),
    
    # OR (95% CI)
    `OR (95% CI)` = sprintf(
      "%.3f (%.3f–%.3f)",
      OR,
      CI_low,
      CI_high
    ),
    
    # main p-value
    overall_p_value = round(p_value, 3),
    
    # coefficient-level interaction p-value
    p_interaction = round(p_int_coef, 3)
  )
```

save table
```{r}
readr::write_csv(
  combined_results,
  file = ""
)
```


plotting
```{r}
plots_by_stratifier <- combined_results_test %>%
  split(.$interaction) %>%
  purrr::imap(function(df, strat_name) {
    
    ggplot(
      df,
      aes(
        x = exposure_label,
        y = OR,
        color = stratum,
        group = stratum
      )
    ) +
      
      geom_point(
        position = position_dodge(width = 0.7),
        size = 3
      ) +
      
      geom_errorbar(
        aes(ymin = CI_low, ymax = CI_high),
        width = 0.25,
        position = position_dodge(width = 0.7)
      ) +
      
      geom_hline(
        yintercept = 1,
        linetype = "dashed",
        color = "gray40"
      ) +
      
      scale_y_log10(
        breaks = seq(0.8, 1.4, by = 0.1),
        labels = scales::number_format(accuracy = 0.1)
      ) +
      
      facet_wrap(~ scenario_title) +
      
      scale_color_jama(name = strat_name) +
      
      labs(
        title = paste("Stratified by", strat_name),
        x = "",
        y = "Odds Ratio"
      ) +
      
      theme_minimal(base_size = 19) +
      
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title  = element_text(face = "bold"),
        strip.text  = element_text(size = 19, face = "bold"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        axis.line.y = element_line(color = "gray50", linewidth = 0.5),
        axis.line.x = element_line(color = "gray50", linewidth = 0.5),
        legend.position = "right" 
      )
  })
```

```{r}
purrr::walk(plots_by_stratifier, print)
```



